---
title: "2024 NFL stats"
author: "Harrison Stanton"
date: "2024-09-10"
output: html_document
---

```{r, message=FALSE}
library(tidyverse)
library(dplyr)
library(rsconnect)
library(kableExtra)
library(nflfastR)
library(nflreadr)
```

Right now, some of the names and stats aren't lining up on the first tab when guys play for two teams in one season. I think by getting teams from the pbp doc this should fix this.

Important thing to note: for first tab, I want to separate when a guy plays on different teams. for the second tab, I don't: they should be lumped. For the third tab, it should be back to if a guy played on two teams, he has two entries. 

This is super close to being right. I just want to make sure that if they played on two teams, they have two observations. Darrynton Evans, for example.

LAST THING TO CHECK BEFORE I PUBLISH IT USING 2024 DATA: 
GUYS WHO PLAYED ON BOTH TEAMS SHOW UP FOR TRAP GRAPH AND SHOW UP FOR WEIGHTED OPPS. SCATTERPLOT IS GREAT RN.

9/7
Everything looked good except for some teams started to have 200% showing on the trap bar graph and I have no idea why. Also, filter gsis_id out of the stats df in the bottom of the page. The Michael Carter/two team issues were fixed though

```{r}
column_names <- c("Total_touches", "Total_HVT", "TRAP", "total_fantasy_points", "total_fantasy_points_exp", "games", "ppg", "xppg", "HVT_perc", "HVT_per_game", "Touches_per_game", "weighted_opps", "wo_per_g")
```


#All of Trap Percentage Doc
```{r}
pbp_2024 <- load_pbp(2024) %>%
    # filter(!(is.na(receiving_yards) & !(is.na(receiver_player_id)))) %>%
    filter(is.na(two_point_conv_result)) %>%
    filter(season_type == "REG")
#don't wanna get team from here, wanna get it from pbp_2023, then group_by(gsis_id, team, week)
roster_2024 <- fast_scraper_roster(2024) %>% 
  select(depth_chart_position, full_name, gsis_id, team, pfr_id) %>% 
  filter(depth_chart_position == "RB")
```

```{r}
#this will work once there are two weeks down...
week_list <- pbp_2024 %>%
  select(week) %>%
  distinct(week) %>%
  pull(week)
```


```{r}
pbp_2024 <- pbp_2024 %>%
  mutate(
    gsis_id = case_when(
      is.na(receiver_player_id) & is.na(rusher_player_id) ~ 'NA',
      is.na(receiver_player_id) ~ rusher_player_id,
      TRUE ~ receiver_player_id
    )
  )
```


```{r}
ff_2024 <- load_ff_opportunity(seasons = 2024, stat_type = "weekly", model_version = "latest")
ff_2024 <- ff_2024 %>% 
  filter(week < 19)

week <- toString(max(ff_2024$week))
season <- toString(max(ff_2024$season))



ff_2024 <- ff_2024 %>%
  group_by(player_id) %>%
  mutate(occurrences = n())

ff_2024 <- ff_2024 %>%
  drop_na(player_id) %>%
  group_by(player_id) %>%
  mutate(occurrences = n())


ff_2024 <- ff_2024 %>% rename(gsis_id = player_id)

ff_2024 <- ff_2024 %>% filter(gsis_id %in% roster_2024$gsis_id)

col_to_sum <- c("total_fantasy_points", "total_fantasy_points_exp")
```

```{r}
rb_df <- ff_2024 %>%
  group_by(gsis_id, week) %>%
  summarize_at((col_to_sum), sum)
```

This isn't quite right, but I don't have time to figure it out right now. Just trying to get a team variable to the guys from the pbp dataframe, not from roster_2023, that way if a player plays on two teams his stats will appear for both.

```{r}
tt_all <- pbp_2024 %>%
   group_by(gsis_id, week) %>%
   filter(is.na(two_point_conv_result)) %>%
   filter((!(is.na(receiving_yards) & !(is.na(receiver_player_id)))) | !is.na(rusher_player_id)) %>%
   summarize(
    Total_touches = n(),
    team = first(posteam)
    )

hvt_all <- pbp_2024 %>%
  group_by(gsis_id, week) %>%
  filter((!(is.na(rusher_player_id)) & yardline_100 <= 10) | (!(is.na(receiver_player_id))) & incomplete_pass == 0 & interception == 0) %>%
  summarize(
    High_value_touches = n(),
    team = first(posteam)
  )
```

```{r}
tt_all <- tt_all %>%
  filter(!gsis_id == "NA")

hvt_all <- hvt_all %>%
  filter(!gsis_id == "NA")
```




Looks good for this part... Michael Carter is an example of a guy who actually logged games on two teams 
```{r}
# tt_all <- pbp_2023 %>%
#   group_by(gsis_id, week) %>%
#   filter(is.na(two_point_conv_result)) %>%
#   filter(!(is.na(receiving_yards) & !(is.na(receiver_player_id)))) %>%
#   summarize(
#     Total_touches = n(),
#     team = first(posteam)
#     )
# 
# hvt_all <- pbp_2023 %>%
#   group_by(gsis_id, week) %>%
#   filter((!(is.na(rusher_player_id)) & yardline_100 <= 10) | (!(is.na(receiver_player_id))) & incomplete_pass == 0 & interception == 0) %>%
#   summarize(
#     High_value_touches = n(),
#     team = first(posteam)
#   )
```

```{r}
teams <- roster_2024 %>%
  distinct(team, .keep_all = TRUE) %>%
  arrange(team) %>%
  pull(team)

new_entries <- c("RB1", "RB2", "RB3")
```



#High value touches section
```{r}
hvt_all <- hvt_all %>%
  group_by(gsis_id, week) %>%
  mutate(Total_HVT = sum(High_value_touches)) %>%
  ungroup()



hvt_all <- hvt_all %>%
  distinct(gsis_id, week, .keep_all = TRUE)


hvt_all <- hvt_all %>%
  select(gsis_id, week, Total_HVT, team)

#hvt_all <- hvt_all %>% rename(gsis_id = receiver_player_id)
```

#High value + total touch data frame
```{r}
hvt_tt_df <- left_join(roster_2024 %>% select(depth_chart_position, gsis_id, full_name, pfr_id), tt_all, by = "gsis_id")
```

```{r}
hvt_tt_df <- left_join(hvt_tt_df, hvt_all %>% select(gsis_id, week, Total_HVT), by = c("gsis_id", "week"))
```

Might need to drop the NA in name and stuff like that, not in total_hvt especially if they played... that's important
Maybe add in a column for this if Total_touches is showing up as NA
Probably drop NA for week is more ideal... that would be if they didn't even appear on special teams 
```{r}
hvt_tt_df$Total_HVT <- ifelse(is.na(hvt_tt_df$Total_HVT), 0, hvt_tt_df$Total_HVT)
hvt_tt_df$Total_touches <- ifelse(is.na(hvt_tt_df$Total_touches), 0, hvt_tt_df$Total_touches)
```

#Need to check this out in different doc
```{r}
#this is comparing guys from roster_2023 to snap_counts to find pfr_ids that didn't automatically show up on roster 2023
#will have to do a section like this in 2024 iteration as well
#list of guys to add eventually
#gonna need to add Jonathon Brooks once he plays, but rn I don't have a pfr_id for him
#Kimani vidal
#Dylan laube
#Blake Watson
roster_2024 <- roster_2024 %>%
  mutate(pfr_id =
         case_when(
           full_name == "Derrick Gore" ~ "GoreDe00",
           full_name == "Patrick Taylor" ~ "TaylPa01",
           full_name == "Tyler Goodson" ~ "GoodTy00",
           full_name == "Jashaun Corbin" ~ "CorbJa01",
           full_name == "Deneric Prince" ~ "PrinDe00",
           full_name == "Chris Brooks" ~ "BrooCh00",
           full_name == "Xazavian Valladay" ~ "VallXa00",
           full_name == "Ray Davis" ~ "DaviRa04",
           full_name == "Trey Benson" ~ "BensTr01",
           full_name == "Carson Steele" ~ "SteeCa00",
           full_name == "Will Shipley" ~ "ShipWi00",
           full_name == "British Brooks" ~ "BrooBr01",
           full_name == "Sione Vaki" ~ "VakiSi00",
           full_name == "Blake Corum" ~ "CoruBl00",
           full_name == "Tyreik McAllister" ~ "McAlTy00",
           full_name == "Isaac Guerendo" ~ "GuerIs00",
           full_name == "Braelon Allen" ~ "AlleBr05",
           full_name == "Bucky Irving" ~ "IrviBu00",
           full_name == "Isaiah Davis" ~ "DaviIs00",
           full_name == "Colson Yankoff" ~ "YankCo00",
           full_name == "Tyrone Tracy Jr." ~"TracTy00",
           full_name == "Audric Estimé" ~ "EstiAu00",
           full_name == "Isaiah Davis" ~ "DaviIs00",
           full_name == "MarShawn Lloyd" ~ "LloyMa00",
           full_name == "Jaylen Wright" ~ "WrigJa04",
           full_name == "Tyler Badie" ~ 	"BadiTy00"
         )
  )
```

#Additions on 8/6
```{r}
#need to add rush_receiver_id to pbp_2023
pbp_2024 <- pbp_2024 %>%
  mutate(
    gsis_id = case_when(
      is.na(receiver_player_id) & is.na(rusher_player_id) ~ 'NA',
      is.na(receiver_player_id) ~ rusher_player_id,
      TRUE ~ receiver_player_id
    )
  )

#think we're good to go here... fixed the NA problems, added 0s instead 
#gonna need this saved in my app

data_table_for_app <- pbp_2024 %>%
  group_by(gsis_id, week) %>%
  summarise(ATT = sum(!is.na(rusher_player_id)),
            Rush_yards = sum(yards_gained[!is.na(rusher_player_id)], na.rm = TRUE),
            "Y/A" = if_else(ATT == 0, 0, round(Rush_yards / ATT, 1)),
            Rush_TD = sum(touchdown[!is.na((rusher_player_id))]),
            REC = sum((!is.na(receiver_player_id)) & complete_pass == 1),
            TGT = sum(!is.na(receiver_player_id)),
            Rec_yards = sum(yards_gained[!is.na(receiver_player_id) & complete_pass == 1]),
            #kinda funny there was a pick 6 on a mccaffrey target that screwed up this number
            "Y/R" = if_else(REC == 0, 0, round(Rec_yards / REC, 1)),
            Rec_TD = sum(touchdown[!is.na(receiver_player_id) & complete_pass == 1]),
            team = first(posteam))
```







#Accurate games played
```{r}
snap_counts <- load_snap_counts(
  seasons = most_recent_season(),
  file_type = getOption("nflreadr.prefer", default = "rds")
)

snap_counts <- snap_counts %>%
  filter(position == "RB")
```




```{r}
snap_counts <- snap_counts %>%
  filter(game_type == "REG") %>%
  group_by(pfr_player_id)

games_df <- snap_counts %>%
  select(player, pfr_player_id, week)

#now I should be able to count games played pretty easily... just count rows for each player after filtering
games_df <- games_df %>%
  distinct(player, week, .keep_all = TRUE)
```

```{r}
rb_df <- left_join(rb_df, roster_2024, by = "gsis_id")
```

```{r}
games_df <- games_df %>% 
  rename(pfr_id = pfr_player_id)
```

```{r}
#this looks goood... prob wanna double check tho. I spot checked Ty Chandler bc there are supposed to be all the weeks appearing except for 8 and 9 and it lookedd good on this 
rb_df <- left_join(rb_df, games_df, by = c("pfr_id", "week"))
```

```{r}
rb_df <- rb_df %>%
  distinct(gsis_id, week, .keep_all = TRUE)
```


```{r}
#need to check this out and make sure the game log is accurate. I think it is, but might need to change this to full_join instead of left_join if it's not 
#just deleted total_fp and total_xfp from the select statement 
hvt_tt_df <- left_join(hvt_tt_df, rb_df %>% select(gsis_id, total_fantasy_points, total_fantasy_points_exp, week), by = c("gsis_id", "week"))
```


#Additions as of 8/5

```{r}
#gonna have to change the names of this once it's 2024 data... just to keep this less confusing
# roster_2023 <- load_rosters(2023) %>%
#   filter(depth_chart_position == "RB")
pbp_2024_rbs <- load_pbp(2024)
```




#rusher_receiver_id
```{r}
#second attempt
pbp_2024_rbs <- pbp_2024_rbs %>%
  mutate(
    gsis_id = case_when(
      is.na(receiver_player_id) & is.na(rusher_player_id) ~ 'NA',
      is.na(receiver_player_id) ~ rusher_player_id,
      TRUE ~ receiver_player_id
    )
  )





#This is correct as of 8/6. Don't change this.
pbp_2024_rbs <- pbp_2024_rbs %>%
  filter(play_type == "run" | play_type == "pass") %>%
  filter(season_type == "REG") %>%
  filter(is.na(two_point_conv_result)) %>%
  filter(gsis_id %in% roster_2024$gsis_id)
```

```{r}
result <- pbp_2024_rbs %>%
  group_by(gsis_id, week) %>%
  summarize(
    rz_targets = sum(yardline_100 < 20 & receiver_player_id == gsis_id, na.rm = TRUE),
    rz_carries = sum(yardline_100 < 20 & rusher_player_id == gsis_id, na.rm = TRUE),
    nrz_carries = sum(yardline_100 >= 20 & rusher_player_id == gsis_id, na.rm = TRUE),
    nrz_targets = sum(yardline_100 >= 20 & receiver_player_id == gsis_id, na.rm = TRUE)
  )



result <- result %>%
  mutate(weighted_opps = 1.3*rz_carries + 2.25*rz_targets + 0.48*nrz_carries + 1.43*nrz_targets)
```

```{r}
hvt_tt_df <- left_join(hvt_tt_df, result %>% select(gsis_id, weighted_opps, week), by = c("gsis_id", "week"))
```

```{r}
#Creating Rank, joining it to the new, expanded hvt df 
# Original data frame
# should be good to count games using this bc the week number comes from games df which comes from snap_counts 
#going to have to change this games filter... don't know what makes sense. > half team games? something like that makes more sense to me tbh, but also if a guy like CMC gets hurt, I still think he needs to be on the list though. 
result <- hvt_tt_df %>%
  group_by(gsis_id) %>%
  summarize(total_fantasy_points = sum(total_fantasy_points),
            games = n(), 
            full_name = first(full_name)) %>%
  mutate(ppg = total_fantasy_points / games) %>%
  #filter(games >= 8) %>%
  arrange(desc(ppg))

# Calculate ranks
rank_bins <- cut(seq(1, nrow(result)),
                 breaks = c(0, 13, 25, 37, nrow(result)),
                 labels = c("RB1", "RB2", "RB3", "Other"),
                 right = FALSE)

# Create a temporary data frame with ranks
result <- result %>%
  mutate(Rank = as.character(rank_bins))


hvt_tt_df <- left_join(hvt_tt_df, result %>% select(gsis_id, Rank), by = "gsis_id")



hvt_tt_df <- hvt_tt_df %>%
  mutate(Rank = ifelse(is.na(Rank), "Other", Rank))
```

```{r}
data_table_for_app <- left_join(data_table_for_app, hvt_tt_df %>% select(gsis_id, Rank, full_name, week), by = c("gsis_id", "week")) %>%
  filter(!is.na(full_name))
```


#stats_df
```{r}
stats_df <- data_table_for_app %>%
  group_by(gsis_id, team) %>%
  summarize(ATT = sum(ATT),
            Rush_yards = sum(Rush_yards),
            Rush_TD = sum(Rush_TD),
            REC = sum(REC),
            TGT = sum(TGT),
            Rec_yards = sum(Rec_yards),
            Rec_TD = sum(Rec_TD),
            Rank = first(Rank),
            #I think this line might be a problem too
            #need to make guys with two teams two different observations 
            team = first(team),
            full_name = first(full_name)) %>%
  mutate("Y/A" = if_else(ATT == 0, 0, round(Rush_yards / ATT, 1)),
         "Y/R" = if_else(REC == 0, 0, round(Rec_yards / REC, 1)))
```


#trap df
```{r}
trap_df <- hvt_tt_df %>%
  group_by(gsis_id, team) %>%
  summarize(
    total_fantasy_points = sum(total_fantasy_points, na.rm = TRUE),
    total_fantasy_points_exp = sum(total_fantasy_points_exp, na.rm = TRUE),
    Total_touches = sum(Total_touches, na.rm = TRUE),
    Total_HVT = sum(Total_HVT, na.rm = TRUE),
    weighted_opps = sum(weighted_opps, na.rm = TRUE),
    games = n(),
    #gonna have to change this line... if there are two teams a dude plays on, he needs to appear on both... aka
    #create another observation. I have to look into this... this is the last thing I need to change though
    team = first(team),
    full_name = first(full_name),
    Rank = first(Rank)
  ) %>%
  mutate(
    TRAP = (Total_touches - Total_HVT) / Total_touches,
    HVT_perc = 1 - TRAP,
    ppg = total_fantasy_points / games,
    xppg = total_fantasy_points_exp / games,
    "WO/G" = weighted_opps / games,
    diff = ppg - `WO/G`
  ) %>%
  pivot_longer(
    cols = c(TRAP, HVT_perc),
    names_to = "Touch_type",
    values_to = "percentage"
  ) %>%
  #what i just changed
  group_by(gsis_id, team) %>%
  mutate(label_y = cumsum(percentage) - percentage / 2,
         label = if_else(Touch_type == "TRAP", Total_touches - Total_HVT, Total_HVT))
```


#scatterplot df
```{r}
scatterplot_df <- hvt_tt_df %>%
  #filter(week >= input$lower & week <= input$higher) %>%
  group_by(gsis_id) %>%
  summarize(
    total_fantasy_points = sum(total_fantasy_points, na.rm = TRUE),
    total_fantasy_points_exp = sum(total_fantasy_points_exp, na.rm = TRUE),
    Total_touches = sum(Total_touches, na.rm = TRUE),
    Total_HVT = sum(Total_HVT, na.rm = TRUE),
    weighted_opps = sum(weighted_opps, na.rm = TRUE),
    games = n(),
    team = first(team),
    full_name = first(full_name), 
    Rank = first(Rank)
  ) %>%
  mutate(
    TRAP = (Total_touches - Total_HVT) / Total_touches,
    HVT_perc = 1 - TRAP,
    ppg = total_fantasy_points / games,
    xppg = total_fantasy_points_exp / games,
    "WO/G" = weighted_opps / games,
    diff = ppg - `WO/G`,
    Touches_per_game = Total_touches / games,
    HVT_per_game = Total_HVT / games
  )
```


#changing hvt_tt_df structure
```{r}
hvt_tt_df <- hvt_tt_df %>%
  mutate(week = as.integer(week))

#adding the last bit for stats_df
stats_df <- left_join(stats_df, scatterplot_df %>% select(gsis_id, total_fantasy_points, total_fantasy_points_exp, Total_touches, Total_HVT, weighted_opps, games, TRAP, ppg, xppg, "WO/G", diff, Touches_per_game, HVT_per_game), by = "gsis_id")
```

```{r}
trap_df <- trap_df %>%
  ungroup()

stats_df <- stats_df %>%
  ungroup()
```


List of objects I need:
week_list
scatterplot_df
trap_df
stats_df 
hvt_tt_df
column_names
week
season

```{r}
trap_df <- trap_df %>%
  mutate(
    label_y = if_else(percentage == 1, .5, label_y),
    label_y = if_else(percentage == 0, NA_integer_, label_y)
  )
```

#pre week 2
```{r}
# save(scatterplot_df, trap_df, stats_df, hvt_tt_df, column_names, week, season, teams,
#      file = "2024NFL.RData")
```


#post week 2
```{r}
save(week_list, scatterplot_df, trap_df, stats_df, hvt_tt_df, column_names, week, season, teams,
     file = "2024NFL.RData")
```



