---
title: "Draft bucket calculations"
author: "Harrison Stanton"
date: "2025-05-22"
output: html_document
---

```{r, message=FALSE}
library(tidyverse)
library(dplyr)
library(nflfastR)
library(nflreadr)
library(stats)
```

The whole goal of this is to randomly sample each position and bucket to see on average how many games these players were not injured and scored 0 fantasy points. I'm going to have to do manual work, but using some statistical techniques can at least limit that and give an approximation for how many 0s I need to be entering. The point of this is to make my projections for rookies in my other model by avoiding selection bias. Before, my dataset only included guys who played snaps/received targets or things like that, meaning all the bad ones who never made the roster or who were only special teamers were filtered out. This way, I'm giving more accurate projections for what each of these players are. 

Probably added a couple too many 0s (hard to tell from football reference if a guy is injured or if he's ), but I think it's much closer than it was.

```{r}
drafted_rookies <- read.csv("Rookies - Sheet1.csv")
```


```{r, message=FALSE}
draft <- load_draft_picks(seasons = TRUE, file_type = getOption("nflreadr.prefer", default = "rds")) %>%
  filter(season >= 2015) %>%
  filter(position %in% c("WR", "RB", "TE", "QB"))
```


```{r}
temp_WR <- draft %>%
  filter(position == "WR") %>% 
  mutate(
    bucket = 
      case_when(
        pick < 16 ~ 1,
        pick < 33 & pick >= 16 ~ 2,
        pick < 66 & pick >= 33 ~ 3,
        pick < 100 & pick >= 66 ~ 4,
        pick < 170 & pick >= 100 ~ 5,
        TRUE ~ 6
      )
  ) %>%
  arrange(bucket)
```


```{r}
temp_RB <- draft %>%
  filter(position == "RB") %>% 
  mutate(
    bucket = 
      case_when(
        pick < 33 ~ 1,
        pick < 66 & pick >= 33 ~ 2,
        pick < 100 & pick >= 66 ~ 3,
        pick < 140 & pick >= 100 ~ 4,
        pick < 180 & pick >= 140 ~ 5,
        TRUE ~ 6
      )
  )
```


```{r}
temp_QB <- draft %>%
  filter(position == "QB") %>% 
  mutate(
    bucket = 
      case_when(
        pick < 15 ~ 1,
        pick < 32 & pick >= 15 ~ 2,
        pick < 66 & pick >= 32 ~ 3,
        pick < 120 & pick >= 66 ~ 4,
        pick < 180 & pick >= 120 ~ 5,
        pick < 180 & pick >= 150 ~ 6,
        TRUE ~ 7
      )
  )
```


```{r}
temp_TE <- draft %>%
  filter(position == "TE") %>% 
  mutate(
    bucket = 
      case_when(
        pick < 33 ~ 1,
        pick < 68 & pick >= 33 ~ 2,
        pick < 95 & pick >= 68 ~ 3,
        pick < 120 & pick >= 120 ~ 4, 
        pick < 155 & pick >= 120 ~ 5,
        pick < 200 & pick >= 155 ~ 6,
        TRUE ~ 7
      )
  )
```



```{r}
temp <- rbind(temp_QB, temp_RB, temp_WR, temp_TE)
```

```{r}
draft <- inner_join(draft, temp %>% select(pfr_player_id, bucket), by = "pfr_player_id") %>%
  arrange((pfr_player_name))
```

```{r}
draft <- draft %>%
  group_by(bucket, position) %>%
  mutate(
    sampling_id = 1:n()
  )
```


This df (draft) should give me a good way to randomize/figure out who I need to sample/figure out how many games they missed not due to injury. I should be chilling from there. 


Trying to write something that tells me which guys I have to look at. ChatGPT recommended checking every guy's game logs if there were less than 30 observations. That's mad tedius, so I'm going to use rookie database to find dudes that didn't play 100% of the games, so that I can just check those. Should cut time considerably

```{r}
load("RookieDatabase.RData")
```

```{r}
Rookie_database <- Rookie_database %>%
  mutate(
    total_fantasy_points = if_else(position == "TE", total_fantasy_points + 0.5 * receptions, total_fantasy_points)
  )
```

7/3 **IMPORTANT**
This df is rookies who were drafted in rookie drafts. Trying to get a feel for if my bucket size is big enough to get a good sample to project each position and bucket. Think this is a more genuine way to go about it than saying all RBs drafted in round 7 bc a lot of those guys were drafted to play special teams. 

Lowkey way happier with this than what I had. Do think I need to rearrange the buckets though. TE and QB buckets need to get shrunk tbh. 
```{r}
Rookie_database %>%
  filter(full_name %in% drafted_rookies$Name) %>%
  distinct(player_id, .keep_all = TRUE) %>%
  group_by(bucket, position) %>%
  summarise(
    number = n()
  )
```

BUCKET REDISTRIBUTION.
TE bucket 2 is gone, combine with TE bucket 3. QB bucket 3 is gone, combine with QB4, 5 and 6.
TE bucket 5-8 needs to get collapsed down into one.
WR buckets look good tbh.
RB buckets 4 and 5 should probably be redistributed, 6 needs to be combined with 5. 


Adding buckets to Rookie_database
```{r}
# Rookie_database <- inner_join(Rookie_database %>% filter(position %in% c("WR", "RB", "QB", "TE")), temp %>% select(gsis_id, bucket) %>% rename(player_id = gsis_id) , by = "player_id")
```



2021 nfl shifted to 18 weeks 
if((season %in% c("2015", "2016", "2017", "2018", "2019", "2020") & games < 16) | (season %in% c("2021", "2022", "2023", "2024", "2025") & games < 17)), FALSE, TRUE)
Adding full_season, a boolean that tells whether a player played in every game that season
```{r}
Rookie_database <- Rookie_database %>%
  group_by(player_id) %>%
  mutate(
    #number of games played in rookie year
    games = n(),
    #whether someone played in all the games
    full_season = if_else((season %in% c("2015", "2016", "2017", "2018", "2019", "2020") & games < 16) | (season %in% c("2021", "2022", "2023", "2024", "2025") & games < 17), FALSE, TRUE)
  ) 
  
```

```{r}
#all guys who played a full season (whether 16 or 17 games) their rookie year
#not going to need to check these guys for 0s in my random sampling
not_to_check <- Rookie_database %>% 
  filter(full_season == TRUE)

#guys who didn't play a full 16 or 17 games their rookie year. Going to have to check these guys for reasons why.
ppl_to_check <- anti_join(draft, not_to_check %>% rename(pfr_player_name = full_name), by = "pfr_player_name")
```

I think I can feel pretty confident in the first two buckets for RBs, WRs that there won't be 0s. I almost don't think it's necessary to check that. For QBs, I almost don't think I'm going to add 0s in either. I think I should just tell ppl if they're not going to play (like Dillon Gabriel week 1), don't put them into your projection. If they're going to start, then put them in. 

So basically, that leaves me checking everything round 2 or later for RBs and WRs, and all TE stuff. 

Should be able to use this list (ppl_to_check) to decide which ppl I even need to check to see if I need to add 0s. Should make this way easier for when I'm doing this on massive scale/have way more ppl who take real zeros, like the round 7 RBs or something. 

To check:
all buckets of TEs
3-6 WR (bucket 2 is back of the first round, think that should be fine or at least not worth the effort)
2-6 RB

RB bucket 2 - review all
RB bucket 3 - review all
RB bucket 4 - ?
RB bucket 5 - ?
RB bucket 6 - ?

WR bucket 2 - all
WR bucket 3 - ?
WR bucket 4 - ?
WR bucket 5 - ?
WR bucket 6 - ?
```{r}
temp_RB %>%
  distinct(pfr_player_id, .keep_all = TRUE) %>%
  group_by(bucket) %>%
  summarise(
    count = n()
  )
```



#df to view for players to check
```{r}
Rookie_database %>%
  #checking for drafted in rookie drafts
  filter(full_name %in% drafted_rookies$Name) %>%
  distinct(player_id, .keep_all = TRUE) %>%
  #checking by position and bucket
  filter(position == "TE" & bucket == 7) %>%
  #checking for didn't play the full season
  filter(player_id %in% ppl_to_check$gsis_id) %>%
  view()
```

Stopped at Hunter Long

For this df I'm creating, looked at way too many RBs, including dudes who were not getting drafted in rookie drafts. Going to have to filter out those names, kind of like I did earlier in this workbook, like at line 138
It'll have to be something like:
google_sheet %>%
filter(full_name %in% drafted_rookies$full_name)
just to get rid of all of that extra data

Need to add a check to make sure there aren't ppl like Brian Thomas Jr., who has a suffix to the end of their name. Just need to make sure some ppl don't fall through the cracks. could probably code that or manually check, but gotta do something.

Probably just count that the number of distinct players in merged data I've made and in adp data i put in here is the same. Then, if it's different, use anti join to see which names i have to adjust.

Found this on the last guy, absolutely tragic, but football reference has injury reports by team by year. Way better than doing what I did, would've saved so much time

Done with manually finding data though