---
output:
  pdf_document: default
  html_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(scales)
library(RSocrata) # for reading in data
library(lubridate) # for dealing with date-time data
```

# Background

We're going to explore a large data set or traffic crashes to learn about what factors are connected with injuries. We will use data from the city of [Chicago's open data portal](https://data.cityofchicago.org/Transportation/Traffic-Crashes-Crashes/85ca-t3if/about_data). (This activity is derived from a [blog post](https://juliasilge.com/blog/chicago-traffic-model/) by Julia Silge)

```{r}
years_ago <- mdy("01/01/2022") # data from last 2 years. May take time to load!
crash_url <- glue::glue("https://data.cityofchicago.org/Transportation/Traffic-Crashes-Crashes/85ca-t3if?$where=CRASH_DATE >= '{years_ago}'")
crash_raw <- as_tibble(read.socrata(crash_url)) # a new way to read in data, don't worry about it!
```

This dataset is pretty crazy! Take a look at it in the viewer, and then let's do some data munging to get it into a nicer form. 

  -create a variable called `injuries` which indicates if the crash involved injuries or not.
  -create an unknown category for missing `report_type`s
  -decide which other variables to keep
  
```{r}
crash <- crash_raw %>%
  arrange(desc(crash_date)) %>%
  transmute(
    injuries = as.factor(if_else(injuries_total > 0, "injuries", "none")),
    crash_record_id, crash_date, weather_condition, lighting_condition, first_crash_type, roadway_surface_cond, crash_type, prim_contributory_cause, crash_month,
    latitude, longitude
  )
crash_raw <- crash_raw %>% mutate(injuries = ifelse(injuries_total > 0, "injuries", "none"))
```


## Exploratory Data Analysis

Here's a few questions to get you started. 

1. Take a look at crashes by latitude and longitude, colored by injuries. What do you notice?

2. What are the most common contributing factors to a crash?

3. How do crashes vary month by month? Compare crashes by month in 2022 to 2023.

4. Are crashes more likely to cause injuries when it is rainy and dark? Use the variables `weather_condition` and `lighting_condition` to explore.

5. Choose a question you want to explore, and create an appropriate visual.

```{r}
ggplot(crash, aes(x = longitude, y = latitude, color = injuries)) +
  geom_point() +
  xlab("Latitude") +
  ylab("Longitude") +
  xlim(-88, -87.5) +
  ylim(41.6, 42.1)
  ggtitle("Longitude vs. Latitude")
```

```{r}
crash_filtered <- crash %>% filter(!(prim_contributory_cause %in% c('NOT APPLICABLE', 'UNABLE TO DETERMINE')))
frequency_table <- table(crash_filtered$prim_contributory_cause)
top_5_items <- names(sort(frequency_table, decreasing = TRUE)[1:5])
top5_df <- data.frame(top_5_items)

crash_filtered <- crash_filtered %>% filter(prim_contributory_cause %in% c('FAILING TO YIELD RIGHT-OF-WAY', 'FOLLOWING TOO CLOSELY', 'IMPROPER OVERTAKING/PASSING', 'FAILING TO REDUCE SPEED TO AVOID CRASH', 'DRIVING SKILLS/KNOWLEDGE/EXPERIENCE'))

#don't know what's happening right now... is.factor(prim_contributory_cause) is returning false, and it was ordering it backwards
#this is making me lose my mind it's just not right
crash_filtered$prim_contributory_cause <- factor(
  crash_filtered$prim_contributory_cause,
    levels = c(
      'FAILING TO YIELD RIGHT-OF-WAY',
      'FOLLOWING TOO CLOSELY',
      'IMPROPER OVERTAKING/PASSING',
      'FAILING TO REDUCE SPEED TO AVOID CRASH',
      'DRIVING SKILLS/KNOWLEDGE/EXPERIENCE'
    ),
  ordered = TRUE
  )

#crash_filtered <- crash_filtered[order(crash_filtered$prim_contributory_cause), ]

```


```{r}
ggplot(crash_filtered, aes(y = prim_contributory_cause)) +
  geom_bar() +
  ylab("Primary Contributing Cause") +
  xlab("Count") +
  ggtitle("Top 5 Causes of Accidents in Chicago")
```

```{r}
year = year(crash$crash_date)
crash = mutate(crash, year = year(crash_date))
crash_graph <- filter(crash, year == 2023 | year == 2022)
# crash_graph <- crash_graph %>% mutate(crash_month_adjusted = ifelse(year == 2022, crash_graph$crash_month, crash_graph$crash_month + 12))
```

```{r}
crash_graph <- crash_graph %>% 
  mutate(
    crash_year = as.factor(year(crash_date)),
    month_name = month(crash_month, label = TRUE, abbr = TRUE)
  )
crash_graph <- crash_graph %>% drop_na(crash_year)
crash_graph_df <- crash_graph %>%
  group_by(crash_year, month_name) %>%
  summarize(Injury_count = sum(injuries == 'injuries'))
```

```{r}
ggplot(crash_graph %>% filter(year == 2022 | year == 2023), aes(x = crash_month)) +
  geom_bar(stat = "count") +
  geom_text(stat = "count", aes(label = ..count..), vjust = -0.5, size = 3) +
  xlab("Crash Month") +
  facet_wrap(~year) +
  scale_x_continuous(breaks = 1:12, labels = month.abb) +
  xlim(0, 13) +
  ylim(0, 11000) +
  ggtitle("Crashes per month in 2022 and 2023") 
```



##updated question 3 graph
```{r}
# ggplot(crash_graph_df, aes(x = month_name)) +
#   geom_line(data = crash_graph_df %>% filter(crash_year == 2022), aes(y = Injury_count), color = 'red') +
#   geom_line(data = crash_graph_df %>% filter(crash_year == 2023), aes(y = Injury_count), color = 'blue')
#   

```


```{r}
crash_22 <- crash_graph %>% filter(year == 2022)
Count_2022 <- crash_22 %>%
  group_by(crash_month) %>%
  summarise(Count_2022 = n())
crash_23 <- crash_graph %>% filter(year == 2023)
Count_2023 <- crash_23 %>% 
  group_by(crash_month) %>%
  summarise(Count2023 = n())
g3_df <- merge(Count_2022, Count_2023, by = 'crash_month', all = FALSE)
```




##second attempt at question 4
```{r}
crash_graph <- crash_graph %>% mutate(Weather_conditions_2 = ifelse(weather_condition == 'CLEAR', 'Clear', 'Poor'))
crash_graph <- crash_graph %>% mutate(Lighting_Conditions_2 = ifelse(lighting_condition %in% c('DAWN', 'DAYLIGHT', 'DUSK', 'DARKNESS, LIGHTED ROAD'), 'Clear', 'Poor'))
```

##by lighting conditions
```{r}
ggplot(crash_graph, aes(x = injuries, fill = Lighting_Conditions_2)) +
  geom_bar(stat = "count") +
  geom_text(stat = "count", aes(label = ..count..), vjust = -0.5, size = 3) +
  xlab("Injuries") +
  ylab("Count") +
  facet_wrap(~Lighting_Conditions_2) +
  ggtitle("Injuries in Car Crashes by Lighing Conditions")
```





```{r}
ggplot(crash_graph, aes(x = injuries, fill = Weather_conditions_2)) +
  geom_bar(stat = "count") +
  geom_text(stat = "count", aes(label = ..count..), vjust = -0.5, size = 3) +
  xlab("Injuries") +
  ylab("Count") +
  facet_wrap(~Weather_conditions_2) +
  ggtitle("Injuries by Weather Conditions")
```

##graph of conditions by both
```{r}
crash_graph <- crash_graph %>% 
  mutate(Rainy_and_Poor = ifelse((Weather_conditions_2 == 'Poor' & Lighting_Conditions_2 == 'Poor'), 'Rainy and Poor', 'Other'))
```

```{r}
ggplot(crash_graph, aes(x = injuries, fill = Rainy_and_Poor)) +
  geom_bar(stat = "count") +
  geom_text(stat = "count", aes(label = ..count..), vjust = -0.5, size = 3) +
  facet_wrap(~Rainy_and_Poor) +
  ggtitle("Injuries with Rainy and Poor Conditions")
```






##exploring my own question 
What were the most common types of crashes, and did those lead to injuries often?
```{r}
frequency_table_2 <- table(crash_filtered$first_crash_type)
print(frequency_table_2)
graph5_df <- crash_graph %>% filter(first_crash_type %in% c('SIDESWIPE SAME DIRECTION', 'TURNING', 'ANGLE', 'REAR END'))
```


```{r}
ggplot(graph5_df, aes(x = injuries, fill = first_crash_type)) +
  geom_bar(stat = "count") +
  geom_text(stat = "count", aes(label = ..count..), vjust = -0.5, size = 3) +
  facet_wrap(~first_crash_type) +
  ylim(0, 45000) +
  ggtitle("Injuries by crash type")
```




##Creating factor variable for day of week
```{r}
crash_raw <- crash_raw %>% mutate(
  day_of_crash = factor(
    case_when(
      crash_day_of_week == 1 ~ "Sunday",
      crash_day_of_week == 2 ~ "Monday",
      crash_day_of_week == 3 ~ "Tuesday",
      crash_day_of_week == 4 ~ "Wednesday",
      crash_day_of_week == 5 ~ "Thursday",
      crash_day_of_week == 6 ~ "Friday",
      crash_day_of_week == 7 ~ "Saturday",
    ),
    levels = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"),
    ordered = TRUE
  )
)

```

##practice with summarize
```{r}
dec_jan_crash <- crash_raw %>% filter(crash_month == 1 | crash_month == 12)
dec_jan_crash <- dec_jan_crash %>% mutate(month = 
  factor(
    case_when(
      crash_month == 1 ~ "January",
      crash_month == 12 ~ "December",
    ),
    levels = c("January", "December"),
    ordered = TRUE 
  )
)
dec_jan_crash <- dec_jan_crash %>% filter(!(is.na(injuries)))


# dec_jan_crash %>%
#   #not gonna work bc I put day_of_crash on crash_raw, not dec_jan_crash
#   group_by(day_of_crash, injuries) %>%
#   mutate(
#     summarize(
#       Count = n()
#       )
#     )

```


