---
title: "FF experiments"
author: "Harrison Stanton"
date: "2024-02-16"
output: html_document
---

```{r}
library(nflreadr)
library(tidyverse)
library(dplyr)
library(ggrepel)
```

```{r}
to_percentage <- function(x) {
  return(paste0(round(x * 100, 2), "%"))
}
```

It would be interesting to look year by year at the dudes who outperformed. Then, I could see how their points per touch changes, total points changes, as well as number of touches changes. I think this would be helpful to look at in terms of understanding what overperforming expectation means.

Also how do I want to filter the data? Bo Melton is on the list of guys who underperformed, but he was only at like -3.5 points, which could easily be a dropped TD or something like that, based on how few reps he had. 

```{r}
ff_exp <- load_ff_opportunity(seasons = most_recent_season(), stat_type = "weekly", model_version = "latest")
ff_exp <- ff_exp %>%
  group_by(player_id) %>%
  mutate(occurrences = n())

ff_exp_2022 <- load_ff_opportunity(seasons = 2022, stat_type = "weekly", model_version = "latest")
ff_exp_2022 <- ff_exp_2022 %>%
  drop_na(player_id) %>%
  group_by(player_id) %>%
  mutate(occurrences = n())
```

```{r}
ff_exp <- ff_exp %>%
  drop_na(player_id) %>%
  group_by(player_id) %>%
  mutate(occurrences = n())
```


```{r}
#want to get this into season stats, not week by week ones... need to figure out how to sum every column, grouping by player 
#want to add games played too... mutate a game played column using cumsum I think is the best bet... or just a count, and add it to that player's name each time
col_to_sum <- c("total_fantasy_points", "total_fantasy_points_exp")

wr_te_df <- ff_exp %>%  
  filter(position %in% c("WR", "TE")) %>%
  group_by(full_name) %>%
  summarize_at((col_to_sum), sum)

wr_te_2022_df <- ff_exp_2022 %>%  
  filter(position %in% c("WR", "TE")) %>%
  group_by(full_name) %>%
  summarize_at((col_to_sum), sum)


#Think we should probably tqke out dudes who didn't play many games
# wr_te_2022_df <- wr_te_2022_df %>%
#   filter(rec_fantasy_points >= 120 & occurrences >= 12)

#positive means overperformed, negative means underperformed
wr_te_df <- wr_te_df %>% 
  mutate(diff = total_fantasy_points - total_fantasy_points_exp)

wr_te_2022_df <- wr_te_2022_df %>%
  mutate(diff = total_fantasy_points - total_fantasy_points_exp)
```


```{r}
wr_te_df <- wr_te_df %>% mutate(percent_change = to_percentage(total_fantasy_points / total_fantasy_points_exp))
```


```{r}
wr_te_df <- wr_te_df %>%
  mutate(my_team = ifelse(full_name %in% c("A.T. Perry", "DeVonta Smith", "Jaxon Smith-Njigba", "Chigoziem Okonkwo", "Kadarius Toney", "Rashid Shaheed", "Kayshon Boutte", "Sam LaPorta", "Jelani Woods", "Romeo Doubs"), TRUE, FALSE))
```

QB data frame
```{r}
col_to_sum <- c("total_fantasy_points", "total_fantasy_points_exp")

qb_2023_df <- ff_exp %>%  
  filter(position == "QB") %>%
  group_by(full_name) %>%
  summarize_at((col_to_sum), sum)

qb_2023_df <- qb_2023_df %>% 
  mutate(diff = total_fantasy_points - total_fantasy_points_exp)

occurrences <- ff_exp %>%
  filter(position %in% c("QB")) %>%
  group_by(full_name) %>%
  summarize(max(occurrences)) %>%
  drop_na(full_name)

qb_2023_df <- inner_join(occurrences, qb_2023_df, by = 'full_name')

stroud_df <- qb_2023_df %>%
  filter(full_name == "C.J. Stroud")
```



```{r}
occurrences <- ff_exp %>%
  filter(position %in% c("WR", "TE")) %>%
  group_by(full_name) %>%
  summarize(max(occurrences)) %>%
  drop_na(full_name)

occurrences_2022 <- ff_exp_2022 %>%
  filter(position %in% c("WR", "TE")) %>%
  group_by(full_name) %>%
  summarize(max(occurrences)) %>%
  drop_na(full_name)

wr_te_df <- inner_join(wr_te_df, occurrences, by = NULL)
wr_te_df$occurrences <- wr_te_df$`max(occurrences)`
wr_te_df <- wr_te_df[, !names(wr_te_df) == "max(occurrences)"]

wr_te_2022_df <- inner_join(wr_te_2022_df, occurrences, by = NULL)
wr_te_2022_df$occurrences <- wr_te_2022_df$`max(occurrences)`
wr_te_2022_df <- wr_te_2022_df[, !names(wr_te_2022_df) == "max(occurrences)"]
```


```{r}
wr_te_df %>%
  filter(occurrences >= 12) %>%
  filter(total_fantasy_points >= 120) %>%
ggplot(aes(x = total_fantasy_points_exp, y = total_fantasy_points)) +
  geom_point() +
  #scale_color_brewer(palette = "Set1", direction = -1) +
  geom_abline(slope = 1, intercept = 0, color = 'red') +
  geom_label_repel(data = filter(wr_te_df, full_name %in% c("Davante Adams", "Garrett Wilson", "Nico Collins", "Tyreek Hill")), aes(x = total_fantasy_points_exp, y = total_fantasy_points, label = full_name)) +
  ggtitle("Fantasy Points vs. Expected Fantasy Points") +
  xlab("Fantasy Points Expected") +
  ylab("Fantasy Poins") 
```

```{r}
wr_te_2022_df %>%
  filter(occurrences >= 12) %>%
  filter(rec_fantasy_points >= 120) %>%
ggplot(aes(x = rec_fantasy_points_exp, y = rec_fantasy_points)) +
  geom_point() +
  #scale_color_brewer(palette = "Set1", direction = -1) +
  geom_abline(slope = 1, intercept = 0, color = 'green') +
  geom_label_repel(data = filter(wr_te_df, full_name %in% c("George Kittle", "Jaylen Waddle", "Diontae Johnson", "Kyle Pitts")), aes(x = rec_fantasy_points_exp, y = rec_fantasy_points, label = full_name)) +
  ggtitle("Fantasy Points vs. Expected Fantasy Points") +
  xlab("Receiving Fantasy Points Expected") +
  ylab("Receiving Fantasy Poins")
```


```{r}
wr_te_df %>%
  filter(my_team == TRUE) %>%
ggplot(aes(x = full_name, y = diff)) +
  geom_bar(stat = "identity", alpha = 0.7) +
  coord_flip() +
  geom_text(aes(label = rec_fantasy_points, y = diff / 2)) +
  ggtitle("Difference between Fantasy Points and Expected Fantasy Points") +
  ylab("Difference in Reaized vs. Expected")
  
```

```{r}
MJP_team <- c("Sam LaPorta", "Rashid Shaheed", "Kayshon Boutte", "Kadarius Toney", 
              "Jaxon Smith-Njigba", "DeVonta Smith", "Chigoziem Okonkwo", "A.T.Perry", "Jalen Hurts", "Carson Wentz", "Kenny Pickett", "Javonte Williams", 
              "Jaleel McLaughlin", "Ty Chandler", "Marcus Mariota", "Mason Rudolph", "Drew Lock", "Tyler Huntley", "Bailey Zappe", "Hendon Hooker", "Tank Bibsby", "Keaton Mitchell", "Zach Charbonnet", "Romeo Doubs", "Jelani Woods")
```

```{r}
ff_exp <- ff_exp %>%
  mutate(my_team = ifelse(full_name %in% MJP_team, TRUE, FALSE))
```

```{r}
occurrences_2 <- ff_exp %>%
  group_by(full_name) %>%
  summarize(max(occurrences)) %>%
  drop_na(full_name)

my_team <- ff_exp %>%
  select("my_team", "full_name") %>%
  drop_na(full_name)

col_to_sum_2 <- c("total_fantasy_points", "total_fantasy_points_exp", "total_fantasy_points_diff")
reduced_ff_exp <- ff_exp %>%
  group_by(full_name) %>%
  summarize_at((col_to_sum_2), sum)

reduced_player_id <- ff_exp %>%
  group_by(full_name) %>%
  select(player_id)

reduced_player_id <- unique(reduced_player_id)
my_team <- unique(my_team)

reduced_ff_exp <- inner_join(reduced_ff_exp, occurrences_2, by = "full_name")
reduced_ff_exp <- inner_join(reduced_ff_exp, reduced_player_id, by = "full_name")
```




My attempt at getting my whole teams expected vs. realized on the same graph... having trouble with the my_team variable atm... was gonna try to arrange them the same way before inner joining and see if that changed anything
```{r}
# #df for just expected vs. realized for all positions
# reduced_ff_exp %>%
#   #my_team variable didn't transfer over right and I have no idea why...
#   filter(my_team == TRUE) %>%
#   #filter(position == 'QB') %>%
# ggplot(aes(x = full_name, y = total_fantasy_points_diff)) +
#   geom_bar(stat = "identity", alpha = 0.7) +
#   coord_flip() +
#   geom_text(aes(label = total_fantasy_points, y = total_fantasy_points_diff / 2)) +
#   ggtitle("Difference between Fantasy Points and Expected Fantasy Points") +
#   ylab("Difference in Reaized vs. Expected")
  
```


```{r}
qb_2023_df %>%
  ggplot(aes(x = total_fantasy_points_exp, y = total_fantasy_points)) +
    geom_point() +
    #scale_color_brewer(palette = "Set1", direction = -1) +
    #geom_point()
    geom_abline(slope = 1, intercept = 0, color = 'blue') +
    geom_text_repel(data = qb_2023_df %>% filter(full_name %in% c("C.J. Stroud")), aes(x = total_fantasy_points_exp, y = total_fantasy_points, label = full_name)) +
    ggtitle("Fantasy Points vs. Expected Fantasy Points") +
    xlab("Fantasy Points Expected") +
    ylab("Fantasy Points")

```


```{r}
espn_qbr <- load_espn_qbr(league = c("nfl"), seasons = most_recent_season(), summary_type = c("season"))
espn_qbr <- espn_qbr %>%
  filter(season_type == "Regular")
```

