---
title: "Best Ball Win prob/projection"
author: "Harrison Stanton"
date: "2025-01-07"
output: html_document
---


To do list (3/18)
1. Might need to rethink how the optimize and generate functions are done. What if I passed two teams into the generate one? and then the add function was something like add_player(score, position, team) ? I think that might be the best way to do it so I only have to run the optimize and generate functions once 
2. Figure out why I can't use summarize on long_scores_df
  couple of ideas: 1. i didn't ungroup temp 2. test if i can use group_by(player_id) on the normal df 3. retest essentially
```{r, message=FALSE}
library(tidyverse)
library(dplyr)
library(rsconnect)
library(kableExtra)
library(nflfastR)
library(nflreadr)
library(stats)
library(Hmisc)
```

```{r, message=FALSE}
load("All_Players_Database.RData")
load("RookieDatabase.RData")
```

```{r}
ff2 <- load_ff_opportunity(2024)
```

```{r}
ff2 %>% 
  summarize(points = mean(total_fantasy_points), .by = player_id)
```


#adding bucket to/pick to ff
```{r}
Rookie_database <- Rookie_database %>% 
  mutate(
    bucket = 
      case_when(
        pick < 11 ~ 1,
        pick < 22 & pick >= 11 ~ 2,
        pick < 33 & pick >= 22 ~ 3,
        round == 2 ~ 4,
        round == 3 ~ 5,
        round == 4 ~ 6,
        round == 5 ~ 7,
        round == 6 | round == 7 ~ 8
      )
  )
```

#data cleaning
```{r}
ff <- All_Players_Database %>% 
  filter(season == "2023" | season == "2022" | season == "2024") %>%
  group_by(player_id) %>%
  slice_head(n=20) %>%
  ungroup()
```

```{r}
ff <- ff %>%
  group_by(player_id) %>%
  mutate(
    deviation = sd(total_fantasy_points, na.rm = TRUE),
    mean = mean(total_fantasy_points, na.rm = TRUE)
  ) %>%
  ungroup()
```

this is filtering out outlier games... anything above or below 3 standard deviations (thinking of Charbonnet bonkers game, Jauan jennings crazy game, etc. Don't want those factoring into my projection)
```{r}
ff <- ff %>%
  filter((is.rookie == FALSE & total_fantasy_points <= mean + 3 * deviation & total_fantasy_points >= mean - 3 * deviation) | is.rookie == TRUE)
```

Rookie data cleaning
```{r}
Rookie_database <- Rookie_database %>%
  group_by(position, bucket) %>%
  mutate(
    deviation = sd(total_fantasy_points, na.rm = TRUE),
    mean = mean(total_fantasy_points, na.rm = TRUE)
  ) %>%
  ungroup()
```

```{r}
Rookie_database <- Rookie_database %>%
  filter((total_fantasy_points <= mean + 3 * deviation & total_fantasy_points >= mean - 3 * deviation)) %>%
  filter(total_fantasy_points >= 0) %>%
  #added this line so there'd be no double counting once next season starts
  filter(season != "2025")
```



```{r}
temp <- Rookie_database %>%
  distinct(player_id, .keep_all = TRUE) %>%
  select(player_id, round, pick, bucket)
```

I think this should be what I want... ff with round, pick, bucket in it. 
```{r}
ff <- left_join(ff, temp, by = "player_id")
```


Jared's team Week 1
Order: 
1. Joe Burrow 2. Bijan Robinson 3. Chase Brown 4. Mike Evans 5. De'Von Achane 6. Rhamondre Stevenson 7. Alec Pierce 8. Tucker Kraft 9. Breece Hall 10. Geno Smith 11. Ezekiel Elliot 12. Demarcus Robinson 13. Adam Thielen 14. Michael Pittman, 15. Keenan Allen 16. Andrei Iosivas 17. Johnny Mundt 18. Tre Tucker 19. Raheem Mostert 20. Gerald Everett 21. Jake Browning 22. Javon Baker 23. TJ Hockenson 24. Charlie Jones 25. Baker Mayfield 
```{r}
Jared_week1 <- c("00-0036442", "00-0038542", "00-0038597", "00-0031408", "00-0039040", "00-0036875", "00-0037664", "00-0038996", "00-0038120", "00-0030565", "00-0033045", "00-0032775", "00-0030035", "00-0036252", "00-0030279", "00-0038619", "00-0033246", "00-0038563", "00-0031687", "00-0033895", "00-0035100", "00-0039853", "00-0035229", "00-0038576", "00-0034855")
```


00-0035100
This is deleting Jake Browning since Joe Burrow was starting week 1
```{r}
Jared_alt <- c("00-0036442", "00-0038542", "00-0038597", "00-0031408", "00-0039040", "00-0036875", "00-0037664", "00-0038996", "00-0038120", "00-0030565", "00-0033045", "00-0032775", "00-0030035", "00-0036252", "00-0030279", "00-0038619", "00-0033246", "00-0038563", "00-0031687", "00-0033895", "00-0039853", "00-0035229", "00-0038576", "00-0034855")
```












#First test
```{r}
for(i in 1:5000){
  rng <- runif(1, min = 0, max = 1)
  if(rng < .25){
    lower <- unname(player_quantile[1])
    upper <- unname(player_quantile[2])
    lower_percent <- 0
    higher_percent <- .25
  } else if(rng >= .25 & rng < .5){
    lower <- unname(player_quantile[2])
    upper<- unname(player_quantile[3])
      lower_percent <- .25
    higher_percent <- .5
  } else if(rng >=.5 & rng < .75){
    lower <- unname(player_quantile[3])
    upper<- unname(player_quantile[4])
    lower_percent <- .5
    higher_percent <- .75
  } else{
    lower <- unname(player_quantile[4])
    upper <- unname(player_quantile[5])
    lower_percent <- .75
    higher_percent <- 1
  }
  modeled_points <- lower + (rng - lower_percent) * (upper - lower)/(higher_percent-lower_percent)
  scores <- c(scores, modeled_points)
  #probably going to want to include something like this:
  #going to need to be careful with that index too... not going to be i in this case I don't think.
  #player_ids_vector <- ids[[x]]
  #colnames(all_player_scores) <- player_ids_vector
  all_player_scores <- data.frame(scores)
}
#in the last for loop, that's where I'm going to reset scores vector. Need to add it to the df first though
```


This should correctly generate every player's quantile in the dataset. 
Now I just need to mesh the code I have above that gives me one score one time into this. 
```{r}
for(i in 1:length(ids)){
  player <- ff %>%
    filter(posteam == "BUF") %>%
    filter(season == "2023")
  
  player_quantile <- player$total_fantasy_points %>%
    #might need to generalize this b, specifically how many groups, too
    quantile(probs = seq(0,1,.25))
  
   #print(player_quantile)
  }
```

#Log Sim
Every variable will have a 2, that means it's the log version, not the normal
Messed up somewhere in here
```{r}
#stores column names with player's ids
player_ids_vector_2 <- c()
for(i in 1:2){
  #vector containing generated scores for one individual
  individual_scores_2 <- c()
  #df with only the one player's stats
  player_2 <- ff %>%
    filter(player_id == ids[[i]])
  
  #quantile for said one player
  #Need to transform at this step
  player_2$log_fp <- log1p(player_2$total_fantasy_points)
  player_quantile_2 <- player_2$log_fp %>%
    #might need to generalize this b, specifically how many groups, too
    quantile(probs = seq(0,1,.25))
  
      for(j in 1:8500){
        rng_2 <- runif(1, min = 0, max = 1)
        if(rng_2 < .25){
          lower_2 <- unname(player_quantile_2[1])
          upper_2 <- unname(player_quantile_2[2])
          lower_percent_2 <- 0
          higher_percent_2 <- .25
        } else if(rng_2 >= .25 & rng_2 < .5){
          lower_2 <- unname(player_quantile_2[2])
          upper_2 <- unname(player_quantile_2[3])
          lower_percent_2 <- .25
          higher_percent_2 <- .5
        } else if(rng_2 >=.5 & rng_2 < .75){
          lower_2 <- unname(player_quantile_2[3])
          upper_2 <- unname(player_quantile_2[4])
          lower_percent_2 <- .5
          higher_percent_2 <- .75
        } else{
          lower_2 <- unname(player_quantile_2[4])
          upper_2 <- unname(player_quantile_2[5])
          lower_percent_2 <- .75
          higher_percent_2 <- 1
        }
        #self explanatory, but using the quantile and interpolation to take the rng, make it the correct value
        interpolated_log <- lower_2 + (rng_2 - lower_percent_2) * (upper_2 - lower_2)/(higher_percent_2-lower_percent_2)
        #transforming back to fantasy points
        modeled_points_2 <- expm1(interpolated_log)
        individual_scores_2 <- c(individual_scores_2, modeled_points_2)
        
        
      }
     if(i == 1){
         all_player_scores_2 <- data.frame(individual_scores_2)
     } else{
         all_player_scores_2 <- cbind(all_player_scores_2, individual_scores_2)
     }
      player_ids_vector_2 <- c(player_ids_vector_2, ids[[i]])
}
```

```{r}
colnames(all_player_scores_2) <- player_ids_vector_2
```


#Log sim updated structure (2/25)
Didn't even test this, looks fairly close, I got tired and turned in for the night. 
It ran, but I need to test the results and see if they're booty before I make this a function. 
```{r}
log_sim <- function(ff, fantasy_team) {
#stores column names with player's ids
player_ids_vector_2 <- c()
all_player_scores_list_2 <- list()
  ids_2 <- ff %>%
    filter(player_id %in% fantasy_team) %>%
    distinct(player_id) %>%
    pull(player_id)
for(i in 1:length(ids_2)){
  #df with only the one player's stats
  player_2 <- ff %>%
    filter(player_id == ids_2[[i]])
    
  
  #quantile for said one player
  #Need to transform at this step
  player_2$log_fp <- log1p(player_2$total_fantasy_points)
  player_quantile_2 <- player_2$log_fp %>%
    #might need to generalize this b, specifically how many groups, too
    quantile(probs = seq(0,1,.25))
  
    individual_scores_2 <- numeric(12000)
  
      for(j in 1:12000){
        rng_2 <- runif(1, min = 0, max = 1)
        if(rng_2 < .25){
          lower_2 <- unname(player_quantile_2[1])
          upper_2 <- unname(player_quantile_2[2])
          lower_percent_2 <- 0
          higher_percent_2 <- .25
        } else if(rng_2 >= .25 & rng_2 < .5){
          lower_2 <- unname(player_quantile_2[2])
          upper_2 <- unname(player_quantile_2[3])
          lower_percent_2 <- .25
          higher_percent_2 <- .5
        } else if(rng_2 >=.5 & rng_2 < .75){
          lower_2 <- unname(player_quantile_2[3])
          upper_2 <- unname(player_quantile_2[4])
          lower_percent_2 <- .5
          higher_percent_2 <- .75
        } else{
          lower_2 <- unname(player_quantile_2[4])
          upper_2 <- unname(player_quantile_2[5])
          lower_percent_2 <- .75
          higher_percent_2 <- 1
        }
        #include this in case upper and lower quantile equal each other (it was always gonna return 0)
        if(upper_2 == lower_2){
          #runif part just adds a little variation, not a huge deal since i'm running this 7500 times and this is a                 fringe case
          modeled_points_2 <- expm1(lower)
        } else {
          interpolated_log <- lower_2 + (rng_2 - lower_percent_2) * (upper_2 - lower_2)/(higher_percent_2-lower_percent_2)
        #transforming back to fantasy points
        modeled_points_2 <- expm1(interpolated_log)
        }
        individual_scores_2[j] <- modeled_points_2
      }
      all_player_scores_list_2[[i]] <- individual_scores_2
      player_ids_vector_2 <- c(player_ids_vector_2, ids_2[[i]])
}

all_player_scores_2 <- as.data.frame(do.call(cbind, all_player_scores_list_2))

colnames(all_player_scores_2) <- player_ids_vector_2



all_player_scores_2 <- cbind(observation = 1:12000, all_player_scores_2)

long_scores_2_df <- all_player_scores_2 %>%
  pivot_longer(cols = -observation, names_to = "player_id", values_to = "score")

#adding position, name to the df (basically so I can check my work better)
long_scores_2_df <- long_scores_2_df %>%
  left_join(ff %>%
              distinct(player_id, .keep_all = TRUE) %>%
              select(player_id, full_name, position), by = "player_id")

return(long_scores_2_df)
}
```


```{r}
log_averages <- jared %>%
  group_by(player_id, full_name) %>%
  summarize(
    log_avg = mean(score)
  ) %>%
  ungroup()
```

```{r}
linear_averages <- jared %>%
  group_by(player_id) %>%
  summarize(
    lin_avg = mean(score)
  ) %>%
  ungroup()
```



```{r}
colnames(all_player_scores_2) <- player_ids_vector_2
```

```{r}
linear_averages <- c()
for(i in 1:18){
  individual_avg <- all_player_scores %>%
    pull(ids[[i]]) %>%
    mean()
  linear_averages <- c(linear_averages, individual_avg)
}

print(linear_averages)
```

```{r}
bills_averages$linear_averages <- linear_averages
```

```{r}
transformed_averages <- c()
for(i in 1:18){
  individual_avg_2 <- all_player_scores_2 %>%
    pull(ids[[i]]) %>%
    mean()
  transformed_averages <- c(transformed_averages, individual_avg_2)
}

print(transformed_averages)
```

```{r}
bills_averages$linear_averages <- linear_averages
```

```{r}
bills_averages$transformed_averages <- transformed_averages
```




#linear sim 2/18
Guys like Raheem Mostert are kind of the locally accurate thing that Seth Walder was talking about. I think this is a solid model, but that's gonna be a place where my ish falls short is with ppl whose roles have clearly changed. I could probably add a condition where he gets minus touches, kind of like I'm planning to add for Charbonnet. That feels like a lot of manual work for during the season, but I'll work on it though.

Probably only wanna do this for bigger names/handcuffs. Not gonna boost guys like Brenton Strange from 2024, but Charbonnet/Raheem Mostert feels more worth my time.
```{r}
#stores column names with player's ids
player_ids_vector <- c()
all_player_scores_list <- list()
ids <- ff %>%
  filter(player_id %in% Jared_alt) %>%
  distinct(player_id) %>%
  pull(player_id)
for(i in 1:length(Jared_alt)){
  #vector containing generated scores for one individual
  individual_scores <- c()
  #df with only the one player's stats
  player <- ff %>%
    filter(player_id == ids[[i]])
  
  #quantile for said one player
  player_quantile <- player$total_fantasy_points %>%
    #might need to generalize this b, specifically how many groups, too
    quantile(probs = seq(0,1,.25))
  
    individual_scores <- numeric(7500)
    
      for(j in 1:7500){
        rng <- runif(1)
        if(rng < .25){
          lower <- unname(player_quantile[1])
          upper <- unname(player_quantile[2])
          lower_percent <- 0
          higher_percent <- .25
        } else if(rng >= .25 & rng < .5){
          lower <- unname(player_quantile[2])
          upper<- unname(player_quantile[3])
            lower_percent <- .25
          higher_percent <- .5
        } else if(rng >=.5 & rng < .75){
          lower <- unname(player_quantile[3])
          upper <- unname(player_quantile[4])
          lower_percent <- .5
          higher_percent <- .75
        } else{
          lower <- unname(player_quantile[4])
          upper <- unname(player_quantile[5])
          lower_percent <- .75
          higher_percent <- 1
        }
        #include this in case upper and lower quantile equal each other (it was always gonna return 0)
        if(upper == lower){
          #runif part just adds a little variation, not a huge deal since i'm running this 7500 times and this is a                 fringe case
          modeled_points <- lower
        } else {
          modeled_points <- lower + (rng - lower_percent) * (upper - lower)/(higher_percent-lower_percent)
        }
        individual_scores[j] <- modeled_points
      }
      all_player_scores_list[[i]] <- individual_scores
      player_ids_vector <- c(player_ids_vector, ids[[i]])
}

all_player_scores <- as.data.frame(do.call(cbind, all_player_scores_list))

colnames(all_player_scores) <- player_ids_vector



all_player_scores <- cbind(observation = 1:7500, all_player_scores)

long_scores_df <- all_player_scores %>%
  pivot_longer(cols = -observation, names_to = "player_id", values_to = "score")

#adding position, name to the df (basically so I can check my work better)
long_scores_df <- long_scores_df %>%
  left_join(ff %>%
              distinct(player_id, .keep_all = TRUE) %>%
              select(player_id, full_name, position))
```


#Linear sim function 2/25
This looks good. I'm honestly pretty happy with this overall. I think I might need to make the fantasy_team input a little different (a df with name and player_id so that building the shiny app will be easier)
Also might want to increase number of observations too. Other than that I'm happy
```{r}
linear_sim <- function(ff, fantasy_team) {
#stores column names with player's ids
player_ids_vector <- c()
all_player_scores_list <- list()
ids <- ff %>%
  filter(player_id %in% fantasy_team) %>%
  distinct(player_id) %>%
  pull(player_id)
for(i in 1:length(ids)){
  #vector containing generated scores for one individual
  individual_scores <- c()
  #df with only the one player's stats
  player <- ff %>%
    filter(player_id == ids[[i]])
  
  #quantile for said one player
  player_quantile <- player$total_fantasy_points %>%
    #might need to generalize this b, specifically how many groups, too
    quantile(probs = seq(0,1,.25))
  
    individual_scores <- numeric(12000)
    
      for(j in 1:12000){
        rng <- runif(1)
        if(rng < .25){
          lower <- unname(player_quantile[1])
          upper <- unname(player_quantile[2])
          lower_percent <- 0
          higher_percent <- .25
        } else if(rng >= .25 & rng < .5){
          lower <- unname(player_quantile[2])
          upper<- unname(player_quantile[3])
            lower_percent <- .25
          higher_percent <- .5
        } else if(rng >=.5 & rng < .75){
          lower <- unname(player_quantile[3])
          upper <- unname(player_quantile[4])
          lower_percent <- .5
          higher_percent <- .75
        } else{
          lower <- unname(player_quantile[4])
          upper <- unname(player_quantile[5])
          lower_percent <- .75
          higher_percent <- 1
        }
        #include this in case upper and lower quantile equal each other (it was always gonna return 0)
        if(upper == lower){
          #runif part just adds a little variation, not a huge deal since i'm running this 7500 times and this is a                 fringe case
          modeled_points <- lower
        } else {
          modeled_points <- lower + (rng - lower_percent) * (upper - lower)/(higher_percent-lower_percent)
        }
        individual_scores[j] <- modeled_points
      }
      all_player_scores_list[[i]] <- individual_scores
      player_ids_vector <- c(player_ids_vector, ids[[i]])
}

all_player_scores <- as.data.frame(do.call(cbind, all_player_scores_list))

colnames(all_player_scores) <- player_ids_vector



all_player_scores <- cbind(observation = 1:12000, all_player_scores)

long_scores_df <- all_player_scores %>%
  pivot_longer(cols = -observation, names_to = "player_id", values_to = "score")

#adding position, name to the df (basically so I can check my work better)
long_scores_df <- long_scores_df %>%
  left_join(ff %>%
              distinct(player_id, .keep_all = TRUE) %>%
              select(player_id, full_name, position))
return(long_scores_df)
}
```

#Rookie quantile work
I'm way too exhausted to work on this right now. I think what I need to do is add buckets to my rookie database. Something like for QBs, WRs and RBs high first, mid first, low first, 2nd, third, 4th and 5th, 6th and 7th as buckets. TEs I think the first round will have to be 2 buckets. 

I should really just start small and then work up. Something like just trying to build a weighted quantile for a specific player like Malik Nabers. Then as ish starts going more, build it out bigger

I think wtd.quantile helped me think of a better way to do this. Definitely am gonna want to add the rookie draft capital to ff though. But it takes two vectors, the scores and then the weights (equal length). So what i think I should do is if they're a rookie, take the scores they already have, then go to Rookie database, filter if they're in the same bucket, then capture all those scores, then build out the weighted quantile. So I will have to add which bucket they're in to ff. I think the easiest way might just be to make a temp df that has the player id, round, pick and bucket, and then just join then and save it as ff. Everyone who isn't a rookie should have NAs in this column. I'll have to add buckets to AllRookies as well



Changed this to 7 for testing purposes. Want to see if the average for nabers using this method of sim is close to Sleeper's projection for him. I'm not necessarily trying to be better on a player projection front, but just have it be close bc my team projection method is way better. 
```{r}
nabers <- ff %>%
  filter(season == "2024") %>%
  filter(player_id == "00-0039337") %>%
  filter(week <= 7)
```

Trying to take Nabers' position and bucket from ff
```{r}
player_position <- ff %>%
  filter(player_id == "00-0039337") %>%
  distinct(player_id, .keep_all = TRUE) %>%
  pull(position)
```

```{r}
player_bucket <- ff %>%
  filter(player_id == "00-0039337") %>%
  distinct(player_id, .keep_all = TRUE) %>%
  pull(bucket)
```


```{r}
WR_bucket1_scores <- Rookie_database %>% 
  filter(position == player_position & bucket == player_bucket) %>%
  pull(total_fantasy_points)
```

```{r}
nabers_scores <- nabers %>%
  pull(total_fantasy_points)
```

Nabers' scores are still first, which is great. 
```{r}
length(c(nabers_scores, WR_bucket1_scores))
```

```{r}
#player + bucket
tot_num_scores <- length(WR_bucket1_scores)
#number of individual's scores
player_num_scores <- length(nabers_scores)
#number of just bucket player's scores
bucket_num_scores <- length(WR_bucket1_scores)
```

Trying to make this so that it weights the player scores as close to 50% with the bucket scores. Don't even know if 50 is the best number tbh, i'm gonna have to mess around with that 
```{r}
player_weights <- bucket_num_scores / player_num_scores
combo_weights <- c(rep(player_weights, player_num_scores), rep(1, bucket_num_scores))
```


Need to look into this. need to find an easy way to make a vector that has weights in it, need to see if the weights thing is even an argument in this b
Weights worked, gonna need to filter out outliers for this ish too... JaMarr chase had a 55.6 as a rookie lol

This is so horny!!!
```{r}
wtd.quantile(c(nabers_scores, WR_bucket1_scores), weights = combo_weights)
```

#3/11 generalized rookie quantile
Need to try adding this to linear sim function as an if else and see what happens. I feel pretty good about this though tbh.
```{r}
#Nabers, Ladd, Daniels, Tyrone Tracey, Bowers, Keon, Rome, AD Mitchell
id_rookies <- c("00-0039337", "00-0039915", "00-0039910", "00-0039384", "00-0039338", "00-0039901", "00-0039919", "00-0039890")
id_short_rookies <- c("00-0039337", "00-0039915")
```

```{r}
#if the player pulled from ids is a rookie
#if((player %>% distinct(player_id, .keep_all = TRUE))$is.rookie == TRUE){

  player_position <- ff %>%
  filter(player_id == "00-0039337") %>%
  distinct(player_id, .keep_all = TRUE) %>%
  pull(position)
  
  player_bucket <- ff %>%
  filter(player_id == "00-0039337") %>%
  distinct(player_id, .keep_all = TRUE) %>%
  pull(bucket)
  
  #scores from the same position and bucket as the rookie we're looking at
  position_bucket_scores <- Rookie_database %>% 
  filter(position == player_position & bucket == player_bucket) %>%
  pull(total_fantasy_points)
  
  player_scores <- player %>%
  pull(total_fantasy_points)
  
  # #player + bucket
  # tot_num_scores <- length(WR_bucket1_scores)
  #number of individual's scores
  player_num_scores <- length(player_scores)
  #number of just bucket player's scores
  bucket_num_scores <- length(position_bucket_scores)
  
  #how rookie's scores are being weighted
  player_weights <- bucket_num_scores / player_num_scores
  #combo of rookie's scores weights and bucket's scores weights
  combo_weights <- c(rep(player_weights, player_num_scores), rep(1, bucket_num_scores))
  
  player_quantile <- wtd.quantile(c(player_scores, position_bucket_scores), weights = combo_weights)
  
  print(player_quantile)
#}

```


**Need to test how this looks**
Idk if this is gonna be the best right now. I think individual player's scores might need more weight than it's getting rn from the looks of it. Also need to test how this is going to work as time goes on... so like after week 3 is my projection dog crap

How do I test this? How does this fit into my linear sim? 

#multiple rookie test, 3/15
I should figure out how to make this loop next. If i can do that/pull the position and bucket for each guy, I think I can do this pretty well in terms of finding at least a kind of efficient way to make a table to check what I should be weighing these things at. 

This is close, not quite right. it's only saving for ladd rn, not the first guy in the df. need to take a look at this tonight. 

I think the problem is that I'm redefining all player scores list and player ids vector again in the loop when it's already done outside of it. I think that just got messed up when I combined the Nabers one player example with the two players example.
```{r}
player_ids_vector <- c()
all_player_scores_list <- list()
ids <- ff %>%
  filter(player_id %in% id_short_rookies) %>%
  distinct(player_id) %>%
  pull(player_id)
for(i in 1:length(ids)){
  #vector containing generated scores for one individual
  individual_scores <- c()
  #df with only the one player's stats
  player <- ff %>%
    filter(player_id == ids[[i]]) %>%
    arrange(desc(season))
  
  if((player %>% distinct(player_id, .keep_all = TRUE))$is.rookie == TRUE){

 
  player_position <- ff %>%
  filter(player_id == ids[[i]]) %>%
  distinct(player_id, .keep_all = TRUE) %>%
  pull(position)
  
  player_bucket <- ff %>%
  filter(player_id == ids[[i]]) %>%
  distinct(player_id, .keep_all = TRUE) %>%
  pull(bucket)
  
  #scores from the same position and bucket as the rookie we're looking at
  position_bucket_scores <- Rookie_database %>% 
  filter(position == player_position & bucket == player_bucket) %>%
  pull(total_fantasy_points)
  
  player_scores <- player %>%
  pull(total_fantasy_points)
  

  #number of individual's scores
  player_num_scores <- length(player_scores)
  #number of just bucket player's scores
  bucket_num_scores <- length(position_bucket_scores)
  
  #how rookie's scores are being weighted
  player_weights <- bucket_num_scores / player_num_scores
  #combo of rookie's scores weights and bucket's scores weights
  combo_weights <- c(rep(player_weights, player_num_scores), rep(1, bucket_num_scores))
  
  player_quantile <- wtd.quantile(c(player_scores, position_bucket_scores), weights = combo_weights)
  }
  
  individual_scores <- numeric(12000)
    
  for(j in 1:12000){
          rng <- runif(1)
          if(rng < .25){
            lower <- unname(player_quantile[1])
            upper <- unname(player_quantile[2])
            lower_percent <- 0
            higher_percent <- .25
          } else if(rng >= .25 & rng < .5){
            lower <- unname(player_quantile[2])
            upper<- unname(player_quantile[3])
              lower_percent <- .25
            higher_percent <- .5
          } else if(rng >=.5 & rng < .75){
            lower <- unname(player_quantile[3])
            upper <- unname(player_quantile[4])
            lower_percent <- .5
            higher_percent <- .75
          } else{
            lower <- unname(player_quantile[4])
            upper <- unname(player_quantile[5])
            lower_percent <- .75
            higher_percent <- 1
          }
          #include this in case upper and lower quantile equal each other (it was always gonna return 0)
          if(upper == lower){
            modeled_points <- lower
          } else {
            modeled_points <- lower + (rng - lower_percent) * (upper - lower)/(higher_percent-lower_percent)
          }
          individual_scores[j] <- modeled_points
        }
         all_player_scores_list[[i]] <- individual_scores
         player_ids_vector <- c(player_ids_vector, ids[[i]])
}
  
   all_player_scores <- as.data.frame(do.call(cbind, all_player_scores_list))
   
   colnames(all_player_scores) <- player_ids_vector
  
  
  
  all_player_scores <- cbind(observation = 1:12000, all_player_scores)
  
  long_scores_df <- all_player_scores %>%
    pivot_longer(cols = -observation, names_to = "player_id", values_to = "score")

  #adding position, name to the df (basically so I can check my work better)
  # long_scores_df <- long_scores_df %>%
  #   ungroup() %>%
  #   left_join(ff %>%
  #               distinct(player_id, .keep_all = TRUE) %>%
  #               select(player_id, full_name, position), by = "player_id") 
```


This line is pissing me off so much. I have no idea what is wrong. My structures are right, i have no NAs. Seriously have no clue why this is happening, even ChatGPT is confused af
```{r}
linear_averages <- long_scores_df %>%
  group_by(player_id) %>%
  summarize(
    lin_avg = mean(score)
  ) %>%
  ungroup()
```


Right now, that looks like a pretty decent number. might've come in a little hot, not sure. 
Should really test guys like Ladd McConkey and Marv later in the season too, see what happens with them. I think that would give me some good outlier cases and might help me make the weights better, but this is a really cool start though.

3/10 - This is good stuff. 
Probably need a more concrete way to determine what my weights are/if my projection was bad. I should probably go through and do this for each week for a couple guys tbh, preferrably all different point levels/positions, and see if there's a number that works best. Maybe start with like 20-80, if one seems better, try numbers in those ranges, stuff like that? 

Want to store his position and bucket in two variables so that I can go back into rookie database and pull the scores of players from that bucket. How do I pull them by player instead of by index? 
Think this is closer to being able to work with now... Still gotta think about what i wanna do with it
I wanna pull the position and bucket for each player, then go to all rookie database, filter the ppl in the same position and bucket, and then combine that with Malik Nabers' scores. Then use wtd.quantile to make a rookie projection. Too tired to think straight about this rn. 



#optimize function
My boy Chet Gupta clutched tf up here
```{r}
optimize_team <- function(long_scores_df) {
optimized_team <- long_scores_df %>%
  arrange(observation, desc(score)) %>%
  group_by(observation) %>%
  mutate(
    qb_rank = cumsum(position == "QB"),
    rb_rank = cumsum(position == "RB"),
    wr_rank = cumsum(position == "WR"),
    te_rank = cumsum(position == "TE"),
  ) %>%
  filter(
    (position == "QB" & qb_rank <= 1) |
    (position == "RB" & rb_rank <= 2) |
    (position == "WR" & wr_rank <= 2) |
    (position == "TE" & te_rank <= 1)
  ) %>%
  ungroup()

leftovers <- long_scores_df %>%
  anti_join(optimized_team, by = c("observation", "player_id"))

flex_players <- leftovers %>%
  filter(position %in% c("RB", "WR", "TE")) %>%
  arrange(observation, desc(score)) %>%
  group_by(observation) %>%
  slice_head(n = 2) %>%
  ungroup()

second_leftovers <- anti_join(long_scores_df, optimized_team, by = c("observation", "player_id"))

second_leftovers <- anti_join(second_leftovers, flex_players, by = c("observation", "player_id"))

superflexes <- second_leftovers %>%
  arrange(observation, desc(score)) %>% 
  group_by(observation) %>%
  slice_head(n=1) %>%
  ungroup()

optimized_team <- rbind(optimized_team %>% select(observation, player_id, score, full_name, position), flex_players, superflexes) %>%
  arrange(observation)

return(optimized_team)
}
```

```{r}
Jared_sim <- linear_sim(ff, Jared_week1)
```

```{r}
Jared_optimal <- optimize_team(Jared_sim)
```

#compare tab
I think when I actually put this on my app, I should just truncate the scores. I don't want it to be so obvious that it changes all the time, and I don't think it would change the experience for anyone. Could pretty easily do the same thing for the max and min projection in the box under the graph too. 
```{r}
compare_tab <- test %>%
  group_by(observation) %>%
  summarize(
    score = sum(score)
  ) %>%
  ungroup()

mean(compare_tab$score)
```

```{r}
compare_tab2 <- Jared_bb2 %>%
  group_by(observation) %>%
  summarize(
    score = sum(score)
  ) %>%
  ungroup()

mean(compare_tab2$score)
```


#adding players who already played
```{r}
add_player <- function(long_scores_df, user_score, user_position){
  observation <- c(1:7500)
  temp <- tibble(player_id = 1, score = user_score, full_name = "Chair", position = user_position)
  temp <- cbind(observation, temp)
  temp <- rbind(long_scores_df, temp)
  temp <- temp %>%
    arrange(observation, desc(score))
  return(temp)
}
```



Side note: of course the linear projection fit super well. I was using data that was from the future to predict what happens. That doesn't mean not knowing the weeks it will perform better.
I want to test 4 things:
1. Linear sim
2. Log sim
3. Linear sim with 3 SDs filtered out - leaning linear sim instead. Still need to actually prove this (don't think i'm gonna... too much work rn to filter out everyone's ish.), but it's looking like this one is just better. Runs faster, projected closer for Jared's week 1
4. Log sim with 3 SDs filtered out

Then want to try this again for the rookies. 


Next steps for my project:
0. I think I'm gonna need to fix way outlier numbers. Or even just the upper quadrant for guys. Like Josh Allen's projected ppg is just too high. I think it's just going to be tough for ppl with mega outlier scores, but idk how to solve for this. First thought is fitting it to a x^2 function instead of just a line, but idk how to do that off the top of my head. 
  The Josh Allen projection isn't as bad as I thought. This data is factoring out Week 18 where he took one snap to keep
  the snaps streak, so his ppg is pretty close to the projection. I'm going to test the log based way too to see if 
  that's better, otherwise I'm fine with the linear interpolation. 
1. make this work for 25 person chunks of ff (going to run into problems when I want to generate two teams... think I'm
just going to make another function and then put all the variable names as 2 lol)
2. figure out how to make this work so that if a user passes the name of a player, I can pull their ids into the "ids" vector that currently contains all bills players
3. Big thing is going to be figuring out how to optimize the scoring to match best ball scoring. The biggest problem I foresee is needing to have each player's positions and everything like that saved too.
  a. One thought I had earlier about this was going through and optimizing each lineup one at a time, saving the total scores to a team1 and team2 vector, and then at the end making that win prob and projected points (I think the win prob and projected points will be the easiest part of the project tbh)
4. making the database for rookies
5. making this a shiny app so me and leaguemates can interact with it

Biggest problems i foresee with my projection:
1. handcuff players 
  what happens when KW3 goes down and my charb projection stays the same? obviously that's not right
      Best guess right now: add a term - fp/touch * normal dist centered around touches vacqted. going to have to
      manually input for guys (i think i can guesstimate how a backfield is gonna be split than actually have 
      a mathematical reasoning behind it)
        - need to test using 2023 data on 2024 season. 
2. rookies
  what happens when 3 games into the season, it's obvious that BTJ and Ladd are better than their quadrant but I'm still
  grouping by their draft capital? Is there some way I can weight recent production more than the past, especially for 
  young guys? 
    Hmisc:wtd.quantile() - weight the persons scores equally to the scores of rookies in the same bucket. I think this
    will be a much more effective way to estimate their scores, but I'm gonna try it using 2023 data on 2024
    season to see what the best method is

**Important**
When I'm telling everyone how to use this, need to make sure that they don't add a bunch of ppl who are hurt/backup QBs. If they're not gonna play, just don't add them. I don't have time to update them all every week, but I'll update handcuff running backs and stuff like that where I can

Going to need a remove player from list function, and also a clear_all function for the shiny app.
