---
title: "Quantile Regression attempt"
author: "Harrison Stanton"
date: "2026-04-29"
output: html_document
---

```{r, message=FALSE}
library(tidyverse)
library(dplyr)
library(rsconnect)
library(kableExtra)
library(nflfastR)
library(nflreadr)
library(slider)
```

```{r}
load("ModelData.RData")
```

Something that I should think about in the future is whether to attempt to project target share and rush share. I left it out in this iteration of the model because I would struggle mightily trying to project these things at the start of the season, to the point where I don't even think it would be helpful. However, going forward, I could see how that would be really helpful (like when a player gets injured), and once I become more cracked, I might look to add something like this to the model too. 

Chat with Claude labeled "Quantile regression models for fantasy football scoring" will be INCREDIBLY HELPFUL for this. Continue to use this as a learning opportunity, dive into the under the hood stuff. I don't want to walk out the other side of this with just a cool model, I want to know how and why it works and be able to defend what I did. 


Run these tests:
1. look at josh jacobs from 2025. Make sure injury stuff looks right, windows calculate properly
2. check out one more TE, WR, QB, make sure scoring adds up/sanity check that
3. Run checks below. If all of those check out, just move on. There's nothing more I can do.


# Check fantasy point distribution makes sense
Besides playoff games, this looks good.
- Playoff games are fixed as of 5/11
```{r}
model_data %>%
  filter(position == "TE") %>%
  summarise(
    min = min(total_fantasy_points, na.rm = TRUE),
    median = median(total_fantasy_points, na.rm = TRUE),
    mean = mean(total_fantasy_points, na.rm = TRUE),
    max = max(total_fantasy_points, na.rm = TRUE)
  )
```

Good
```{r}
# Check for duplicate player-game rows
model_data %>%
  count(gsis_id, season, week) %>%
  filter(n > 1)
```

```{r}
# Check rolling stats reset correctly between seasons
model_data %>%
  filter(gsis_id == "00-0035700", week == 1) %>%
  select(season, week, total_16g_n, carries_16g_avg, ypc_16g)
```

```{r}
#rolling stats still calculate correctly even with injuries
model_data %>%
  filter(gsis_id == "00-0040734") %>%
  arrange(desc(season), desc(week)) %>%
  view()
```

Still to do as of 5/11
Last check, just look at one guy per position, make sure fant numbers line up, stats look right, betting lines look ok, etc.



Beginning analysis of data (probably should do point above this first, but I'm tired of checking data for tonight)
```{r}
# Step 1: Filter to RBs and build chronological game index per player
rb_data <- model_data %>%
  filter(position == "RB") %>%
  arrange(gsis_id, season, week) %>%
  group_by(gsis_id) %>%
  mutate(game_num = row_number()) %>%
  ungroup()
```

```{r}
# Step 2: Compute correlation at each lag across all RBs
max_lag <- 70

lag_cors <- map_dfr(1:max_lag, function(lag) {
  rb_data %>%
    group_by(gsis_id) %>%
    mutate(lagged_pts = dplyr::lag(total_fantasy_points, lag)) %>%
    ungroup() %>%
    filter(!is.na(lagged_pts)) %>%
    summarise(
      lag = lag,
      correlation = cor(total_fantasy_points, lagged_pts, use = "complete.obs")
    )
})
```

```{r}
# Step 3: Plot
ggplot(lag_cors, aes(x = lag, y = correlation)) +
  geom_point() +
  geom_line() +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
  scale_x_continuous(breaks = seq(0, 30, by = 2)) +
  labs(
    title = "Autocorrelation of RB Fantasy Points by Lag (Games)",
    x = "Lag (games in the past)",
    y = "Correlation with current game"
  ) +
  theme_minimal()
```


```{r}
# Step 2: Autocorrelation on carries instead of fantasy points
max_lag <- 70

lag_cors_carries <- map_dfr(1:max_lag, function(lag) {
  rb_data %>%
    group_by(gsis_id) %>%
    mutate(lagged_carries = dplyr::lag(rush_attempt, lag)) %>%
    ungroup() %>%
    filter(!is.na(lagged_carries)) %>%
    summarise(
      lag = lag,
      correlation = cor(rush_attempt, lagged_carries, use = "complete.obs")
    )
})
```

```{r}
# Step 3: Plot both on the same chart for easy comparison
lag_cors_fantasy <- lag_cors %>% mutate(variable = "Fantasy Points")
lag_cors_carries <- lag_cors_carries %>% mutate(variable = "Carries")

bind_rows(lag_cors_fantasy, lag_cors_carries) %>%
  ggplot(aes(x = lag, y = correlation, color = variable)) +
  geom_point() +
  geom_line() +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
  scale_x_continuous(breaks = seq(0, 30, by = 2)) +
  labs(
    title = "Autocorrelation: Fantasy Points vs Carries by Lag",
    x = "Lag (games in the past)",
    y = "Correlation with current game",
    color = NULL
  ) +
  theme_minimal()
```

