---
title: "Player databases"
author: "Harrison Stanton"
date: "2025-02-11"
output: html_document
---

Updated as of 6/30
```{r, message=FALSE}
library(tidyverse)
library(dplyr)
library(rsconnect)
library(kableExtra)
library(nflfastR)
library(nflreadr)
```


```{r}
rookies <- load_draft_picks() %>%
  filter(position %in% c("QB", "WR", "TE", "RB")) %>%
  filter(season >= 2015) %>%
  select(season, round, pick, team, gsis_id, pfr_player_name, position)
```



```{r}
rookies <- rookies %>%
  mutate(
    season_player_id = paste(season, gsis_id, sep = " ")
  )
```


```{r}
for(i in 2015:2024){
    ff <- load_ff_opportunity(seasons = i)
    
    ff <- ff %>%
      mutate(
        season_player_id = paste(season, player_id, sep = " ")
               )
    
    ff <- ff %>%
      filter(season_player_id %in% rookies$season_player_id) %>%
      select(season, week, posteam, game_id, player_id, season_player_id, full_name, position, pass_yards_gained, pass_touchdown, pass_interception,pass_two_point_conv, receptions, rec_attempt, rec_yards_gained, rec_touchdown, rec_fumble_lost, rec_two_point_conv, rush_attempt, rush_yards_gained, rush_touchdown, rush_two_point_conv, rush_fumble_lost, total_fantasy_points)
    
    ff<- ff %>%
      mutate(
        half_ppg = 0.5 * receptions + .04 * pass_yards_gained + .1 * (rec_yards_gained + rush_yards_gained) + 4 * pass_touchdown + 6 * (rush_touchdown + rec_touchdown) + 2 * (pass_two_point_conv + rush_two_point_conv + rec_two_point_conv) + -2 * (pass_interception + rec_fumble_lost + rush_fumble_lost)
    )
    
    ff <- ff %>%
      filter(!is.na(player_id))
    
    if(i == 2015){
      Rookie_database <- data.frame(ff)
    } else{
      Rookie_database <- rbind(Rookie_database, ff)
    }
}
```


```{r}
Rookie_database <- Rookie_database %>%
  rename(gsis_id = player_id)
```


```{r}
Rookie_database <- inner_join(Rookie_database, rookies %>% select(gsis_id, round, pick), by = "gsis_id")
```

```{r}
Rookie_database <- Rookie_database %>%
  rename(player_id = gsis_id)
```

Extra 0 weeks
```{r, message=FALSE}
extra_zeros <- read_csv("Missing weeks Rookie data - Sheet1.csv") %>%
  filter(!is.na(season)) 
extra_zeros$week <- NA
```

Adding healthy scratches/zero fp games
```{r}
Rookie_database <- Rookie_database %>%
  rbind(extra_zeros)
```

#Adding buckets to df without temp so i don't get error when rejoining
UPDATED 7/22
I still need to redo these buckets. Somewhere (draftbuckets.rmd? i redid these in a better way, need to shrink bucket sizes, especially for QB and TE)

Going to have to redo these as well once I look at the numbers/position breakdowns for rookies actually drafted in dynasty drafts
```{r}
Rookie_database <- Rookie_database %>%
    mutate(
    bucket = 
      case_when(
        position == "WR" & pick < 16 ~ 1,
        position == "WR" & pick < 33 & pick >= 16 ~ 2,
        position == "WR" & pick < 66 & pick >= 33 ~ 3,
        position == "WR" & pick < 100 & pick >= 66 ~ 4,
        position == "WR" & pick < 170 & pick >= 100 ~ 5,
        position == "WR" & TRUE ~ 6,
        position == "RB" & pick < 33 ~ 1,
        position == "RB" & pick < 66 & pick >= 33 ~ 2,
        position == "RB" & pick < 100 & pick >= 66 ~ 3,
        position == "RB" & pick < 180 & pick >= 100 ~ 4,
        position == "RB" & TRUE ~ 5,
        position == "TE" & pick < 33 ~ 1,
        position == "TE" & pick < 68 & pick >= 33 ~ 3,
        position == "TE" & pick < 95 & pick >= 68 ~ 4,
        position == "TE" & pick < 120 & pick >= 120 ~ 5, 
        position == "TE" & pick < 155 & pick >= 120 ~ 6,
        position == "TE" & pick < 200 & pick >= 155 ~ 7,
        position == "TE" & TRUE ~ 8,
        position == "QB" & pick < 15 ~ 1,
        position == "QB" & pick < 65 & pick >= 11 ~ 2,
        position == "QB" & pick < 120 & pick >= 66 ~ 3,
        position == "QB" & pick < 180 & pick >= 120 ~ 4,
        position == "QB" & pick < 180 & pick >= 150 ~ 5,
        position == "QB" & TRUE ~ 6
      )
    )
```

#Adding buckets to df
All code from before adding in extra zeros
```{r}
# temp1 <- Rookie_database %>%
#   filter(position == "WR") %>% 
#   mutate(
#     bucket = 
#       case_when(
#         pick < 16 ~ 1,
#         pick < 33 & pick >= 16 ~ 2,
#         pick < 66 & pick >= 33 ~ 3,
#         pick < 100 & pick >= 66 ~ 4,
#         pick < 170 & pick >= 100 ~ 5,
#         TRUE ~ 6
#       )
#   )
```


```{r}
# temp2 <- Rookie_database %>%
#   filter(position == "RB") %>% 
#   mutate(
#     bucket = 
#       case_when(
#         pick < 33 ~ 1,
#         pick < 66 & pick >= 33 ~ 2,
#         pick < 100 & pick >= 66 ~ 3,
#         pick < 180 & pick >= 100 ~ 4,
#         TRUE ~ 5
#       )
#   )
```


```{r}
# temp3 <- Rookie_database %>%
#   filter(position == "QB") %>% 
#   mutate(
#     bucket = 
#       case_when(
#         pick < 15 ~ 1,
#         pick < 65 & pick >= 11 ~ 2,
#         pick < 120 & pick >= 66 ~ 3,
#         pick < 180 & pick >= 120 ~ 4,
#         pick < 180 & pick >= 150 ~ 5,
#         TRUE ~ 6
#       )
#   )
```


```{r}
# temp4 <- Rookie_database %>%
#   filter(position == "TE") %>% 
#   mutate(
#     bucket = 
#       case_when(
#         pick < 33 ~ 1,
#         pick < 68 & pick >= 33 ~ 3,
#         pick < 95 & pick >= 68 ~ 4,
#         pick < 120 & pick >= 120 ~ 5, 
#         pick < 155 & pick >= 120 ~ 6,
#         pick < 200 & pick >= 155 ~ 7,
#         TRUE ~ 8
#       )
#   )
```

```{r}
# temp <- rbind(temp1, temp2, temp3, temp4)
```



I can't use week here since all of my week numbers are NA in my google doc... and i saved over rookie_database. That should be easy to do tho, just unjoin
```{r}
# Rookie_database <- left_join(Rookie_database, temp %>% select(player_id, week, bucket), by = c("week", "player_id")) %>%
#   filter(!is.na(full_name))
```

```{r}
save(Rookie_database,
     file = "RookieDatabase.RData")
```
