---
title: "TrapPercentage"
author: "Harrison Stanton"
date: "2024-03-25"
output: html_document
---

```{r, message=FALSE}
library(nflfastR)
library(tidyverse)
library(dplyr)
library(nflreadr)
library(knitr)
library(kableExtra)
```

```{r, message=FALSE}
pbp_2023 <- load_pbp(2023)
roster_2023 <- fast_scraper_roster(2023)
```

```{r}
ff_2023 <- load_ff_opportunity(seasons = most_recent_season(), stat_type = "weekly", model_version = "latest")
```

```{r, message=FALSE}
FP_df <- read_csv("RBstatsCopy.csv")
```


```{r}
ff_2023 <- load_ff_opportunity(seasons = most_recent_season(), stat_type = "weekly", model_version = "latest")

ff_2023 <- ff_2023 %>%
  filter(week < 19)

ff_2023 <- ff_2023 %>%
  group_by(player_id) %>%
  mutate(occurrences = n())

ff_2023 <- ff_2023 %>%
  drop_na(player_id) %>%
  group_by(player_id) %>%
  mutate(occurrences = n())

ff_2023 <- ff_2023 %>% rename(gsis_id = player_id)

ff_2023 <- ff_2023 %>% filter(position == "RB")

col_to_sum <- c("total_fantasy_points", "total_fantasy_points_exp")

rb_df <- ff_2023 %>%  
  group_by(gsis_id) %>%
  summarize_at((col_to_sum), sum)
```


```{r}
roster_2023 <- roster_2023 %>% select(position, full_name, gsis_id, team)
roster_2023 <- roster_2023 %>% filter(position == "RB")
```

```{r}
pbp_2023 <- pbp_2023 %>%
  filter(!(is.na(receiving_yards) & !(is.na(receiver_player_id)))) %>%
  filter(is.na(two_point_conv_result)) %>%
  filter(season_type == "REG")
```

```{r, message=FALSE}
tt_all <- pbp_2023 %>%
  group_by(receiver_player_id, rusher_player_id) %>%
  filter(is.na(two_point_conv_result)) %>%
  filter(!(is.na(receiving_yards) & !(is.na(receiver_player_id)))) %>%
  summarize(Carries_or_catches = n())


hvt_all <- pbp_2023 %>%
  group_by(receiver_player_id, rusher_player_id) %>%
  filter((!(is.na(rusher_player_id)) & yardline_100 <= 10) | !(is.na(receiver_player_id))) %>%
  summarize(High_value_touches = n()) 
```

```{r}
teams <- roster_2023 %>%
  distinct(team, .keep_all = TRUE) %>%
  pull(team)

new_entries <- c("RB1", "RB2", "RB3")
teams <- c(teams, new_entries)
print(teams)
```


```{r}
tt_all$receiver_player_id <- ifelse(is.na(tt_all$receiver_player_id), tt_all$rusher_player_id, tt_all$receiver_player_id)

tt_all$rusher_player_id <- ifelse(is.na(tt_all$rusher_player_id), tt_all$receiver_player_id, tt_all$rusher_player_id)

tt_all <- tt_all %>%
  group_by(receiver_player_id) %>%
  mutate(Total_touches = sum(Carries_or_catches)) %>%
  ungroup()
```

```{r}
tt_all <- tt_all %>%
  distinct(receiver_player_id, .keep_all = TRUE)


tt_all <- tt_all %>%
  select(receiver_player_id, Total_touches)

tt_all <- tt_all %>% rename(gsis_id = receiver_player_id)
```

High value touches section:
```{r}
hvt_all$receiver_player_id <- ifelse(is.na(hvt_all$receiver_player_id), hvt_all$rusher_player_id, hvt_all$receiver_player_id)

hvt_all$rusher_player_id <- ifelse(is.na(hvt_all$rusher_player_id), hvt_all$receiver_player_id, hvt_all$rusher_player_id)

hvt_all <- hvt_all %>%
  group_by(receiver_player_id) %>%
  mutate(Total_HVT = sum(High_value_touches)) %>%
  ungroup()
```

```{r}
hvt_all <- hvt_all %>%
  distinct(receiver_player_id, .keep_all = TRUE)


hvt_all <- hvt_all %>%
  select(receiver_player_id, Total_HVT)

hvt_all <- hvt_all %>% rename(gsis_id = receiver_player_id)
```

```{r}
#High value + total touch data frame
hvt_tt_df <- left_join(roster_2023, tt_all, by = "gsis_id")

hvt_tt_df <- left_join(hvt_tt_df, hvt_all, by = "gsis_id")
```

```{r}
hvt_tt_df <- hvt_tt_df %>% filter(!is.na(Total_touches))
```

```{r}
hvt_tt_df$Total_HVT <- ifelse(is.na(hvt_tt_df$Total_HVT), 0, hvt_tt_df$Total_HVT)
```


```{r}
hvt_tt_df <- hvt_tt_df %>%
  mutate(TRAP = (Total_touches - Total_HVT) / Total_touches)
```


```{r}
games_df <-ff_2023 %>%
  filter(position %in% c("RB")) %>%
  group_by(gsis_id) %>%
  select("occurrences", "gsis_id", "season")

games_df <- games_df %>%
  rename(games = occurrences)
```


```{r}
rb_df <- left_join(rb_df, games_df, by = "gsis_id")
rb_df <- rb_df %>%
  distinct(gsis_id, .keep_all = TRUE)
hvt_tt_df <- left_join(hvt_tt_df, rb_df, by = "gsis_id")
```

```{r}
hvt_tt_df <- hvt_tt_df %>%
  mutate(ppg = total_fantasy_points / games,
         xppg = total_fantasy_points_exp / games)

hvt_tt_df <- hvt_tt_df %>%
  mutate(HVT_perc = 1 - TRAP)
```

```{r}
hvt_tt_df <- hvt_tt_df %>% 
  arrange(desc(total_fantasy_points))
```

```{r}
hvt_tt_long_df <- pivot_longer(hvt_tt_df, cols = c(TRAP, HVT_perc),
                          names_to = "Touch_type", values_to = "percentage")

hvt_tt_long_df <- hvt_tt_long_df %>%
  group_by(gsis_id) %>%
  mutate(label_y = cumsum(percentage) - percentage / 2)
```



```{r}
# Create bins for ranking
rank_bins <- cut(seq(1, nrow(hvt_tt_long_df)), 
                 breaks = c(0, 25, 49, 73, nrow(hvt_tt_long_df)), 
                 labels = c("RB1", "RB2", "RB3", "Other"), 
                 right = FALSE)

# Add Rank column to the data frame
hvt_tt_long_df$Rank <- as.character(rank_bins)
hvt_tt_long_df$Rank <- ifelse(is.na(hvt_tt_long_df$Rank), "Other", hvt_tt_long_df$Rank)
```



```{r}
#this is by RB1s... need to figure out a way to do this for RB2, RB3 as well 
#how do I arrange this? 
#what order is this even presenting in?? I'm so confused, I can't even figure out the pattern
hvt_tt_long_df %>%
  arrange(desc(total_fantasy_points)) %>%
  head(24) %>%
ggplot(aes(x = full_name, y = percentage, fill = Touch_type)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  scale_fill_manual(values = c("springgreen4", "brown2"), name = "Touch Type",
                    labels = c("High Value Touches", "Trivial Rushes")) +
  labs(x = "", y = "Percentage") +
  theme(
        plot.title = element_text(size = 24), 
        axis.title = element_text(size = 18), 
        axis.text.x = element_text(size = 14),
        axis.text.y = element_text(size = 14),
        legend.title = element_text(size = 16), 
        legend.text = element_text(size = 14),
      ) +
  ggtitle("Trivial Rush Attempt Percentage (TRAP) for 2023") +
  geom_text(aes(y = label_y, label = scales::percent(percentage, accuracy = .01)), color = 'white') 
  scale_y_continuous(labels = scales::percent)
    
```

```{r}
names <- hvt_tt_long_df %>%
  filter(Rank == "RB1") %>%
  distinct(gsis_id, .keep_all = TRUE) %>%
  pull(full_name)

names <- rev(names)
```


```{r}
hvt_tt_long_df %>%
  filter(Rank == "RB1") %>%
ggplot(aes(x = full_name, y = percentage, fill = Touch_type)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  scale_fill_manual(values = c("springgreen4", "brown2"), name = "Touch Type",
                    labels = c("High Value Touches", "Trivial Rushes")) +
  labs(x = "", y = "Percentage") +
  theme_minimal() +
  scale_x_discrete(limits = names) +
  ggtitle("Trivial Rush Attempt Percentage (TRAP) for 2023") +
  geom_text(aes(y = label_y, label = scales::percent(percentage, accuracy = .01)), color = 'white') +
  # annotate("text", x = 3, y = 18, label = "Source: nflfastR and Fantasy Pros", 
  #          hjust = 0, vjust = 0, size = 4, color = "gray")
  scale_y_continuous(labels = scales::percent) 
  
```

```{r}
teams <- unique(hvt_tt_long_df$team)
sort(teams)
```

```{r}
hvt_tt_long_df %>%
  filter(Rank == "RB1") %>%
  select(full_name, team, Total_touches, Total_HVT, total_fantasy_points, total_fantasy_points_exp, games, ppg, xppg, Touch_type, percentage)
```

```{r}
# Create bins for ranking
rank_bins <- cut(seq(1, nrow(hvt_tt_df)), 
                 breaks = c(0, 13, 25, 37, nrow(hvt_tt_df)), 
                 labels = c("RB1", "RB2", "RB3", "Other"), 
                 right = FALSE)

# Add Rank column to the data frame
hvt_tt_df$Rank <- as.character(rank_bins)
hvt_tt_df$Rank <- ifelse(is.na(hvt_tt_df$Rank), "Other", hvt_tt_df$Rank)
```



```{r}
hvt_tt_df <- hvt_tt_df %>%
  mutate(HVT_per_game = Total_HVT / games,
         Touches_per_game = Total_touches/games)
```

```{r}
FP_df <- FP_df %>%
  mutate(
    other_touches_game = Touches_per_game - HVT_per_game
      )
```


```{r}
FP_df %>%
      ggplot(aes(x = Touches_per_game, y = HVT_per_game)) +
        geom_point() +
        xlab("Touches Per Game") +
        ylab("HVT per game") +
        ggtitle("HVT per game vs. Touches per game") +
        geom_smooth(method = "lm", se = FALSE) +
      theme(
        plot.title = element_text(size = 24),
        axis.title = element_text(size = 18), 
        axis.text.x = element_text(size = 14),
        axis.text.y = element_text(size = 14),
        legend.title = element_text(size = 16), 
        legend.text = element_text(size = 14),
      )
```

```{r}
lm1 <- lm(ppg ~ HVT_per_game, data = FP_df)
slope <- coef(lm1)[2]
intercept <- coef(lm1)[1]
    
print(paste0("y = ", round(slope, 2), "x + ", round(intercept, 2)))
```

```{r}
lm2 <- lm(ppg ~ Touches_per_game, data = FP_df)
slope <- coef(lm2)[2]
intercept <- coef(lm2)[1]
    
print(paste0("y = ", round(slope, 2), "x + ", round(intercept, 2)))
```

```{r}
lm3 <- lm(ppg ~ HVT_per_game + other_touches_game, data = FP_df)
slope <- coef(lm3)[2]
slope2 <- coef(lm3)[3]
intercept <- coef(lm3)[1]

print(paste0("y = ", round(slope, 2), "x1 + ", round(intercept, 2), "x2 + ", round(slope2, 2)))
```


I feel like I should be looking for backfields where 2.5 HVT/game are up for grabs or people are getting that many touches... 2.5 * 3.18 = 7.95, just need some good variance at that point... I could probably even run a linear regression for the carries within the 10 versus the receptions, or find the EV for those two touches

I also want to take a peek at how many HVT/game a team is even producing for their backfields... so group by team, sum HVT, div 17 will tell us how many high value opportunities there were for the backs, and then from there I can learn more about how much I need a team to be using a guy for this to pop, especially with split backfields
```{r}
FP_df %>%
  filter(rank != 'Other') %>%
      ggplot(aes(x = HVT_per_game, y = ppg)) +
        geom_point() +
        xlab("Touches Per Game") +
        ylab("HVT per game") +
        ggtitle("HVT per game vs. Touches per game") +
        geom_smooth(method = "lm", se = FALSE) +
      theme(
        plot.title = element_text(size = 24),
        axis.title = element_text(size = 18), 
        axis.text.x = element_text(size = 14),
        axis.text.y = element_text(size = 14),
        legend.title = element_text(size = 16), 
        legend.text = element_text(size = 14),
      )
```

