---
title: "Weighted Opportunity Data Frame"
author: "Harrison Stanton"
date: "2024-06-26"
output: html_document
---

```{r, message=FALSE}
library(tidyverse)
library(dplyr)
library(nflfastR)
library(nflreadr)
```

```{r, message=FALSE}
ff_2023 <- load_ff_opportunity(2023)
```

```{r}
week <- toString(max(ff_2023$week))
season <- toString(max(ff_2023$season))
```

```{r}
roster_2023 <- load_rosters(2023) %>%
  filter(depth_chart_position == "RB")
pbp_2023_rbs <- load_pbp(2023)
```


#games played done well
```{r}
snap_counts <- load_snap_counts(
  seasons = most_recent_season(),
  file_type = getOption("nflreadr.prefer", default = "rds")
)
# snap_counts <- snap_counts %>% 
#   filter(position == "RB" | position == "FB")
snap_counts <- snap_counts %>%
  filter(position == "RB")

```


```{r}
snap_counts <- snap_counts %>% 
  filter(game_type == "REG") %>%
  group_by(pfr_player_id) %>%
  mutate(games = n())

games_df <- snap_counts %>%
  select(player, pfr_player_id, games)

games_df <- games_df %>%
  distinct(player, .keep_all = TRUE)
```


#rusher_receiver_id
```{r}
#second attempt
pbp_2023_rbs <- pbp_2023_rbs %>%
  mutate(
    rush_receiver_id = case_when(
      is.na(receiver_player_id) & is.na(rusher_player_id) ~ 'NA',
      is.na(receiver_player_id) ~ rusher_player_id,
      TRUE ~ receiver_player_id
    )
  )
```


```{r}
#only pass or run plays from the regular seasonseason
#it seems like this filter (the run or pass) takes care of penalties that were overturned. 
#the ones that stay in the df are ones that were after the play which get added on, like a late hit
pbp_2023_rbs <- pbp_2023_rbs %>%
  filter(play_type == "run" | play_type == "pass") %>%
 # filter(is.na(penalty_player_id)) %>%
  filter(season_type == "REG") %>%
  filter(is.na(two_point_conv_result))
```



```{r}
#this should get us to the point where it's only running backs involved, and it's pass or run plays...
#still gotta check penalties
pbp_2023_rbs <- pbp_2023_rbs %>%
  filter(rush_receiver_id %in% roster_2023$gsis_id)
```


#calculating touches for each RB
```{r}
result <- pbp_2023_rbs %>%
  group_by(rush_receiver_id) %>%
  summarize(
    rz_targets = sum(yardline_100 < 20 & receiver_player_id == rush_receiver_id, na.rm = TRUE),
    rz_carries = sum(yardline_100 < 20 & rusher_player_id == rush_receiver_id, na.rm = TRUE),
    nrz_carries = sum(yardline_100 >= 20 & rusher_player_id == rush_receiver_id, na.rm = TRUE),
    nrz_targets = sum(yardline_100 >= 20 & receiver_player_id == rush_receiver_id, na.rm = TRUE)
  )
```


```{r}
temp_df <- roster_2023 %>%
  select(gsis_id, full_name, team, pfr_id)

temp_df <- temp_df %>%
  rename(rush_receiver_id = gsis_id,
         pfr_player_id = pfr_id)

result_df <-left_join(temp_df, result, by = "rush_receiver_id")

result_df <- result_df %>%
  drop_na(nrz_carries)
```

```{r}
result_df <- left_join(result_df, games_df, by = "pfr_player_id")
```

```{r}
#fixed... same as Ryan Heath's now... love that 
result_df <- result_df %>%
  mutate(weighted_opps = 1.3*rz_carries + 2.25*rz_targets + 0.48*nrz_carries + 1.43*nrz_targets,
         wo_per_g = weighted_opps / games)
```

```{r}
ff_2023 <- ff_2023 %>%
  filter(week < 19)
```

```{r}
ff_2023 <- ff_2023 %>% 
  rename(rush_receiver_id = player_id)
```

```{r}
ff_2023 <- ff_2023 %>% filter(position == "RB")

col_to_sum <- c("total_fantasy_points")

points_df <- ff_2023 %>%  
  group_by(rush_receiver_id) %>%
  summarize_at((col_to_sum), sum)
```

```{r}
result_df <- left_join(result_df, points_df, by = "rush_receiver_id")
```

```{r}
result_df <- result_df %>%
  mutate(ppg = total_fantasy_points / games)
```

```{r}
# Original data frame
result_df <- result_df %>%
  # Arrange by ppg for temporary ranking
  arrange(desc(ppg))

# Filter out players with fewer than 8 games for ranking purposes
result_df <- result_df %>%
  filter(games >= 8) %>%
  arrange(desc(ppg))

# Calculate ranks
rank_bins <- cut(seq(1, nrow(result_df)), 
                 breaks = c(0, 13, 25, 37, nrow(result_df)), 
                 labels = c("RB1", "RB2", "RB3", "Other"), 
                 right = FALSE)

# Create a temporary data frame with ranks
result_df <- result_df %>%
  mutate(Rank = as.character(rank_bins))
```

```{r}
result_df <- result_df %>%
  rename(gsis_id = rush_receiver_id)
```


```{r}
output_wo_df <- result_df %>%
  select(player, team, games, ppg, wo_per_g, Rank)

output_wo_df <- output_wo_df %>%
  drop_na(player)

output_wo_df <- output_wo_df %>%
  filter(player != "Devon Achane")

output_wo_df <- output_wo_df %>%
  mutate(diff = ppg - wo_per_g)
```
`

```{r}
# Original data frame
output_wo_df <- output_wo_df %>%
  # Arrange by ppg for temporary ranking
  arrange(desc(ppg))

# Filter out players with fewer than 8 games for ranking purposes
temp_df <- output_wo_df %>%
  filter(games >= 8) %>%
  arrange(desc(ppg))

# Calculate ranks
rank_bins <- cut(seq(1, nrow(temp_df)), 
                 breaks = c(0, 13, 25, 37, nrow(temp_df)), 
                 labels = c("RB1", "RB2", "RB3", "Other"), 
                 right = FALSE)

# Create a temporary data frame with ranks
temp_df <- temp_df %>%
  mutate(Rank = as.character(rank_bins))

# Merge the ranks back into the original data frame
output_wo_df <- output_wo_df %>%
  left_join(temp_df %>% select(gsis_id, Rank), by = "gsis_id")

# Ensure players with fewer than 8 games are classified as "Other"
output_wo_df$Rank <- ifelse(is.na(output_wo_df$Rank), "Other", output_wo_df$Rank)
```


```{r}
output_wo_df <- output_wo_df %>%
  arrange(desc(wo_per_g))
```









#Weighted Opportunities graphs
I like both of these graphs quite a bit... think they're both helpful visually to see kind of what we're looking for/what is good and what's bad. 
```{r}
output_wo_df %>%
  ggplot(aes(x = wo_per_g, y = ppg)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(
    x = "Weighted Opportunities per game",
    y = "PPG",
    title = "Weighted Opportunities per game vs. PPG"
  )
```

It would be interesting to look at past years for weighted opportunities as well... Ryan Heath's article talks about regressing to "your mean"... want to see what Raheem Mostert/Derrick Henry/guys like that's means actually look like 

How should I calculate career difference in PPG and WO/G?


```{r}
output_wo_df %>%
  filter(games >= 8 & wo_per_g >= 1.5) %>%
  ggplot(aes(y = diff)) +
  geom_boxplot() +
  labs(
    y = "PPG - WO/G",
    title = "Efficiency analysis for 2023 Running Backs"
  ) +
  theme(
    axis.title.x = element_blank(),  # Remove x-axis title
    axis.text.x = element_blank(),   # Remove x-axis text
    axis.ticks.x = element_blank()   # Remove x-axis ticks
  )
```

```{r}
quantiles <- output_wo_df %>%
  filter(games >= 8 & wo_per_g >= 1.5) %>%
  select(diff) %>%
  unlist() %>% # Convert the selected column to a vector
  quantile(probs = c(0.25, 0.75))

# Print the results
quantiles
```


#code for attempting to merge the WO_per_game to the other Df used in the app
```{r}
rb_stats_df <- left_join(rb_stats_df, result_df %>% select(gsis_id, weighted_opps, wo_per_g), by = "gsis_id")
```

